### 2016-05-24 draft1.tex
tex added

\documentclass{amsart}

%     If your article includes graphics, uncomment this command.
\usepackage{graphicx}
\usepackage{amsmath,amsfonts,amssymb,amscd,amsthm,amsbsy,upref}
\usepackage[all]{xy}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{paralist}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{tikz-cd}
\usepackage{pgfplots}
\usepackage{fancyhdr}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{array}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{xca}[theorem]{Exercise}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\numberwithin{equation}{section}

%    Absolute value notation
\newcommand{\abs}[1]{\lvert#1\rvert}

%    Blank box placeholder for figures (to avoid requiring any
%    particular graphics capabilities for printing this document).
\newcommand{\blankbox}[2]{%
  \parbox{\columnwidth}{\centering
%    Set fboxsep to 0 so that the actual size of the box will match the
%    given measurements more closely.
    \setlength{\fboxsep}{0pt}%
    \fbox{\raisebox{0pt}[#2]{\hspace{#1}}}%
  }%
}

\pgfplotsset{compat=1.13}

\begin{document}

\title{Assigning Scouts to Optimal Patrols}
\author{Phil Snyder}
\author{Ian Bruce}
\author{Julio Marco Pineda}

\begin{abstract}
The Scoutmaster wants to find an arrangement of 13 new scouts into two patrols. The cohesiveness and overall quality experience of a patrol is severely affected by adverse relationship and can be improved by existing preferences. Thus, by using anonymous data from scouts and parents, a brute force strategy was employed to find the optimal arrangement of scouts that avoids severe conflict and maximizes existing preferences. An arrangement of 6 and 7 scouts on one patrol respectively was found that satisfies the desired conditions. 
\end{abstract}
\maketitle
\section*{Problem Description}
Every year the Scoutmaster must assign new scouts to groups called patrols. In these patrols, scouts would plan and perform different scouting activities together in which they can form close friends that grow together. However, in recent years, the number of scouts have been increasing at a rate such that it is no longer logistically viable to place every new scout in a single patrol; multiple patrols are needed to organize activities better and to form proper bonds and community between the new recruits. Furthermore, the new scouts are in the fifth or sixth grade who have not fully developed their interpersonal skills. Many can be temperamental and awkward when dealing with adversity, and even some devastating conflicts occur where not only the two fighting scouts create animosity between each other, the rest of the troop alienates the destructive pair. If enough of these adverse interactions occur within a troop, many scouts would decide to quit and lose out in a great experience and community in the Boy Scouts.

The data the scoutmaster collects about adverse pairings and preferences are either privately given by the scouts themselves or by their parents. The possibly destructive interactions are given by the parents of the scouts.Previously, the Scoutmaster was able to determine the perfect patrols manually due to having smaller number of scouts. However, the number of scouts have been increasing yearly to the point that solving this grouping problem cannot be done by hand in a reasonable time. Therefore, we can help our community partner find a better means to solve his troop organization problem. Our goal for this project is to assign boy scouts in different patrols of appropriate size (of 6-8 scouts) such that we maximize the retention rate of the scouts in the program by avoiding severe conflicts while maximizing positive relationships between scouts.

We would like to answer the direct problem our community partner provided of arranging 15 scouts into two troops that avoids severe conflicts and promotes the positive relationships between the scouts. Furthermore, we want to determine if we can build a model to predict these patrol arrangements for any number of scouts so that the Scoutmaster can use this model for future years. Other than improving the overall experience of the scouts and alleviating the burden of the Scoutmaster, this model can possibly be applied to other situations where arranging a number of people of groups where interpersonal dynamics is necessary and important.

\section*{Simplifications}
Ultimately, our goal is to find groups for the scouts that will allow them to grow and be happy in the troop. If we really wanted to fully understand how a group can foster growth in an individual, we would want to consider the astronomical number of aspects about the personalities of each member and how those influence each of the other members given those other members’ astronomically complex personalities. Sociological and psychological literature would need to be studied extensively to learn about these aspects, and experts in the field would most likely be required to collect relevant data for each child, taking several weeks to conduct interviews. At the of the day, given our limited knowledge in this area, we settled with considering a subset of the sociological data between the scouts. We asked each scout which other scouts they liked and which ones they disliked. Now, this is obviously a giant simplification, but the difficulty of gathering more complex information and compiling said information would be too great for the expertise level of anyone involved in solving this problem. Also, this process of gathering data would have to be repeated each year for each new batch of new scouts, and would therefore be a burden on the Scoutmaster to collect a lot of data.

Another simplification is to assume that the quality of a group can be determined by the quality of the relationships between the unique pairs of the group - in other words, the whole is equal to the sum of the parts. For example, consider the case where scouts A, B and C individually like each other when they are alone with one other scout, but don’t like being in a group together. This epiphenomenon won’t be considered; in context of our model, the pairing of these scouts would be heavily favored.

\section*{Mathematical Model}
This problem visually lends itself well to the notion of a mathematical graph. Since there is an inherent sense of relationships between individuals (how much a scout likes another scout), it is appropriate to consider the scouts as vertices in a graph with directed edges representing how much the “head” scout likes the “tail” scout. Our goal, using this language, is to find a partition of the vertices in the graph that satisfies some sort of objective function given size constraints of the partitions. In our problem at hand, we consider a complete graph with 13 vertices and edge weights being either 0, 1, or -1, corresponding to indifference, approval, or disapproval respectively of the “head” scout with respect to the “tail” scout. The Scoutmaster specifically wants two groups out of these 13, and with the size constraints for a patrol considered, our goal becomes finding 6 vertices to be in one partition and the other 7 being in another. The total number of partitions becomes 13 choose 6, which is equivalent to 1,716. This is certainly within the scope of considering total enumeration of the partitions as a reasonable solution to the problem, therefore we will proceed accordingly and implement the calculation using Python and its associated libraries to represent the data in a format that is convenient to work with.

Now, a natural question for us to consider is what objective function would be most appropriate to optimize. A natural answer to this would be to sum the edge weights of all edges that connect vertices that are within partitions. This would make sense if separating enemies is just as important as joining friends, but this just might not be the case. In terms of weighing these factors, we defer to Scoutmaster Bruce’s experience with how past patrols fared with friends and foes in the mix. Therefore, we consider a family of possible objective functions, namely giving real numbered weights to instances of approving relationships and disapproving relationships, and select a few of these objective functions to solve for. We then present the collection of results from different objective functions to the Scoutmaster for him to see which one seems to most reasonably match his preferences.
\section*{Solution of Mathematical Problem}
We represent our directed graph as a 13 by 13 similarity matrix, or, as might be more fitting for the problem, an “affinity” matrix. Like below:%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\footnote{The scouts were enumerated like so: 1. Brandon, 2. Cameron, 3, Christian, 4. Colby, 5. Daniel, 6. Darwin, 7. Evan, 8. Jake, 9. Jordan, 10. Nathan, 11. Patrick, 12. Timmy, 13. Tommy.}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\includegraphics[scale=0.6]{solution1.JPG}

The NA value represents a hard constraint established by a scout’s parent (i.e., under no circumstances are these two to be in the same patrol). In our implementation we convert the NA value to $-\infty$. We designed an objective function that accepts two scouts who are assumed to be in different patrols and calculates the loss associated with such a placement. The loss is defined to be $L(x, y) = \text{max}(-2, x[y] + y[x])$ where $x[y]$ is $x$’s disposition towards $y$ and vise versa. For example, if $x$ dislikes $y$ and $y$ dislikes $x$, the loss is -2. If $x$ likes $y$ but $y$ dislikes $x$ the loss is 0. If $x$ and $y$ both like each other, the loss is 2. In mixed cases where the first scout is indifferent to another, but the other likes or dislikes the first scout, the loss is +1 or -1 respectively. We are trying to minimize the loss, so in applying our loss function over the entire dataset, our algorithm is equally averse to same-patrol mutual enmity as it is to dividing mutual friendship, half as averse to the indifference of one scout and the stronger inclination of another, and indifferent to polar opposite affinities within the same patrol. There is a question as to whether this is an accurate reflection of how individual scouts would view their own “loss” of any of the possible scenarios. As it turns out, our algorithm arrives at a very reasonable solution regardless. 
\section*{Results}
As mentioned earlier, the number of scouts we must partition is small enough to find an optimal solution by brute force. Our algorithm finds a partition of scouts into patrols of sizes 6 and 7 with a total loss of 0 as described in the table below:
\begin{table}[ht]
	\caption{}\label{eqtable}
	\renewcommand\arraystretch{1.5}
	\noindent\[
	\begin{array}{|c|c|}
	\hline
	\textbf{Patrol 1}&\textbf{Patrol 2}\\
	\hline
	\text{Brandon, Cameron, Colby,}&\text{Christian, Daniel, Jake,}\\
	\text{Darwin, Evan, Tommy}&\text{Jordan, Nathan, Patrick, Timmy}\\
	\hline
	\end{array}
	\]
\end{table}
\includegraphics[scale=0.65]{results1.JPG}
Of the undesirable outcomes of such a partition, we can spot a few:
\begin{enumerate}
	\item Tommy and Timmy both like each other and are in separate patrols. But if we move Timmy to Patrol 1 ($P1$), we break up the mutual friendship of Timmy and Nathan in $P2$. Suppose we moved both Timmy and Nathan to $P1$ and Brandon (who dislikes Nathan) to $P2$. But then Brandon loses two friends (Tommy and Cameron), gains one (Jordan), and now neither Nathan nor Timmy are in the same patrol as Daniel, whom they both like. Although, Brandon and Tommy find themselves in the awkward situation of having complete opposite sentiments towards each other. Perhaps this would be a worthwhile trade after all (and a larger, negative weight needs to be placed upon disfavor-favor relationships).

	\item As mentioned in the previous point, Tommy dislikes Brandon. Trading Tommy in $P1$ for Jordan in $P2$ seems promising, But we have overlooked the fact that Colby and Jordan are our NA pair, and cannot be in the same patrol under any circumstances. Trading Colby for Jordan is another option, but Colby will be missed by Darwin, Evan, Cameron and Tommy (In other words, the whole of P1 minus Brandon).
\end{enumerate}
On the whole, though, there are no obvious improvements that can be made to our algorithm’s partition. We find that this is an assignment that could have reasonably been arrived at by Scoutmaster Bruce (granted the group dynamics have not changed since the scouts were surveyed).  Furthermore, the solution took only 20 seconds to arrive at on a single core machine.

\section*{Improvements}
While our algorithm demonstrably works well on smaller groups of 13 people, we do not expect our brute force solution to be tractable on larger groups wherein greater than 2 partitions are required. As a possible scenario, consider a company of 150 people to be divided into fixed-size teams of 10 individuals each. There are $150! / (10!)^{15}$ possible partitions, a number proportional to $10^164$ - meaning we could instead use that computational time to enumerate the number of atoms in the universe… a tredecillion $(10^{78})$ times. Knowing this, perhaps it won’t surprise the reader that this particular graph partitioning problem in general graphs is NP-complete - though there do exist approximation algorithms.%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\footnote{B. W. Kernighan, S. Lin, \textit{An Efficient Heuristic Procedure for Partitioning Graphs}. Bell System Technical Journal. \textbf{49} (1970), 291--307. doi: 10.1002/j.1538-7305.1970.tb01770.x}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
As scout enrollment increases, it becomes necessary to consider scenarios where we have both a large number of scouts in need of assignment and multiple patrols to choose from. Our algorithm could be extended to use brute force to find a globally optimal solution when it is deemed computationally feasible and to use an approximating algorithm to find a locally optimal solution otherwise.

\section*{Conclusions}
We initially struggled to decide on determining an appropriate objective function to partition the scouts to different groups. Follow up interviews with Scoutmaster Gene Bruce allowed us to decide on this function. We learned to rely on his expertise and intuition rather than trying to blindly decide on which types of relationships should be fostered and avoided to weight our objective function. Thus, we learned necessary communication skills and consideration of our community partner's needs.

With a feasibility check of how many solutions possible once the mathematical model was decided using graphs, a brute force strategy was the most accessible method to arrive at a solution. Thus we were able to develop a program that can partition scouts using this strategy. We learned to utilize tools we have at hand right away, and then we can consider other methods later on to reduce the computational time of our algorithm.

\section*{Acknowledgments}
We would like to thank Scoutmaster Gene Bruce for providing us the necessary data and being extremely patient and understanding when discussing the background and problem. We would also like to thank Professor Sarah Billey for the mentorship and guidance she provided throughout this project, and our TA Austin Tran for his comments and suggestions.

\section*{Verification Statement}
(Once we receive our verification statement, we will place it here.)

\bibliographystyle{amsplain}
\begin{thebibliography}{10}

\bibitem {A} B. W. Kernighan, S. Lin, \textit{An Efficient Heuristic Procedure for Partitioning Graphs}. Bell System Technical Journal. \textbf{49} (1970), 291--307. doi: 10.1002/j.1538-7305.1970.tb01770.x

\end{thebibliography}

\end{document}

%------------------------------------------------------------------------------
% End of journal.tex
%------------------------------------------------------------------------------

### 2016-05-24 main.py
My comment

@@ -1,3 +1,5 @@
+# Julio's here!!!
+
 from __future__ import print_function
 import pandas as pd
 import numpy as np

### 2016-05-25 draft1.tex
Random matrix generator
new section comparing objective functions + edits

@@
 \section*{Mathematical Model}
-This problem visually lends itself well to the notion of a mathematical graph.
-Since there is an inherent sense of relationships between individuals (how much a scout likes another scout), it is appropriate to consider the scouts as vertices in a graph with directed edges representing how much the “head” scout likes the “tail” scout.
-Our goal, using this language, is to find a partition of the vertices in the graph that satisfies some sort of objective function given size constraints of the partitions.
-In our problem at hand, we consider a complete graph with 13 vertices and edge weights being either 0, 1, or -1, corresponding to indifference, approval, or disapproval respectively of the “head” scout with respect to the “tail” scout.
-The Scoutmaster specifically wants two groups out of these 13, and with the size constraints for a patrol considered, our goal becomes finding 6 vertices to be in one partition and the other 7 being in another.
-The total number of partitions becomes 13 choose 6, which is equivalent to 1,716.
-This is certainly within the scope of considering total enumeration of the partitions as a reasonable solution to the problem, therefore we will proceed accordingly and implement the calculation using Python and its associated libraries to represent the data in a format that is convenient to work with.
+To model the relationships between scouts, we use a mathematical graph.
+We let the scouts be vertices in our graph.
+A directed edge between Scout A and Scout B has a weight proportional to how much Scout A likes Scout B.
+We would then like to find a partition of the vertices in the graph that minimizes an objective function given constraints on the sizes of subsets generated by our partition.
+This results in a complete graph on 13 vertices having edge weights 0, 1, or -1, corresponding to indifference, approval, or disapproval, respectively.
+Scoutmaster Bruce specifically would like two patrols from the 13 scouts - one of size 6, the other of size 7.
+The total number of partitions given these constrants is $\binom{13}{6} = \binom{13}{7} = 1716$.
+This number is reasonable enough that we may calculate the total loss of every valid grouping and select the partition with the lowest score.
 
-Now, a natural question for us to consider is what objective function would be most appropriate to optimize.
-A natural answer to this would be to sum the edge weights of all edges that connect vertices that are within partitions.
-This would make sense if separating enemies is just as important as joining friends, but this just might not be the case.
-In terms of weighing these factors, we defer to Scoutmaster Bruce’s experience with how past patrols fared with friends and foes in the mix.
-Therefore, we consider a family of possible objective functions, namely giving real numbered weights to instances of approving relationships and disapproving relationships, and select a few of these objective functions to solve for.
-We then present the collection of results from different objective functions to the Scoutmaster for him to see which one seems to most reasonably match his preferences.
+We measure the goodness (or score) of a partition in six different ways, corresponding to six different objective functions, which we dub MinCut, FriendCut, EnemyCut, AwkwardCut, and HybridCut.
+MinCut is designed to minimize the cut edge weights between partitions (see \textit{Solution of Mathematical Problem}).
+FriendCut is designed to maximize the number of intra-patrol positive edge weights.
+EnemyCut is meant to minimize the number of intra-patrol negative edge weights, Awkward cut ($+i$) is the same as MinCut, but with an additional penalty of $i$ if there are a pair of scouts, Scout A and Scout B, such that Scout A favors Scout B, but Scout B disfavors Scout A.
+Hybrid Cut is a convex combination of MinCut and FriendCut, each with equal weight (that is, half the MinCut loss plus half the FriendCut loss).
+Our three primary objective functions - MinCut, FriendCut, and EnemyCut are given a more rigorous treatment in the next section.
+
 \section*{Solution of Mathematical Problem}
-We represent our directed graph as a 13 by 13 similarity matrix, or, as might be more fitting for the problem, an “affinity” matrix. Like below:%
+We represent our directed graph as a 13 by 13 similarity matrix, or, as might
+be more fitting for the problem, an ``affinity'' matrix.
+Like below\footnote{The scouts were enumerated like so: 1.
+Brandon, 2.
+Cameron, 3, Christian, 4.
+Colby, 5.
+Daniel, 6.
+Darwin, 7.
+Evan, 8.
+Jake, 9.
+Jordan, 10.
+Nathan, 11.
+Patrick, 12.
+Timmy, 13.
+Tommy.}:
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.28]{data}
+\end{figure}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\footnote{The scouts were enumerated like so: 1. Brandon, 2. Cameron, 3, Christian, 4. Colby, 5. Daniel, 6. Darwin, 7. Evan, 8. Jake, 9. Jordan, 10. Nathan, 11. Patrick, 12. Timmy, 13. Tommy.}%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\includegraphics[scale=0.6]{solution1.JPG}
 
-The NA value represents a hard constraint established by a scout’s parent (i.e., under no circumstances are these two to be in the same patrol).
-In our implementation we convert the NA value to $-\infty$.
-We designed an objective function that accepts two scouts who are assumed to be in different patrols and calculates the loss associated with such a placement.
-The loss is defined to be $L(x, y) = \text{max}(-2, x[y] + y[x])$ where $x[y]$ is $x$’s disposition towards $y$ and vise versa.
+The NA value represents a hard constraint established by a scout's parent (i.e., under no circumstances are these two to be in the same patrol).
+We designed a number of objective functions that accept two scouts and calculates the loss associated with such a placement.
+The loss is defined to be $L(x, y) = \text{max}(-2, x[y] + y[x])$ where $x[y]$ is $x$'s disposition towards $y$ and vise versa.
 For example, if $x$ dislikes $y$ and $y$ dislikes $x$, the loss is -2.
@@
 We are trying to minimize the loss, so in applying our loss function over the entire dataset, our algorithm is equally averse to same-patrol mutual enmity as it is to dividing mutual friendship, half as averse to the indifference of one scout and the stronger inclination of another, and indifferent to polar opposite affinities within the same patrol.
-There is a question as to whether this is an accurate reflection of how individual scouts would view their own “loss” of any of the possible scenarios.
+There is a question as to whether this is an accurate reflection of how individual scouts would view their own ``loss'' of any of the possible scenarios.
 As it turns out, our algorithm arrives at a very reasonable solution regardless.
@@
 As mentioned earlier, the number of scouts we must partition is small enough to find an optimal solution by brute force.
-Our algorithm finds a partition of scouts into patrols of sizes 6 and 7 with a total loss of 0 as described in the table below:
-\begin{table}[ht]
+Our algorithm finds a partition of scouts into patrols of sizes 6 and 7 which is optimal with respect to each objective.
+
+\clearpage
+
+\begin{figure}[h]
+\centering
+\begin{tabular}{ |l|l|l|l|l|l| }
+\hline
+\multicolumn{2}{|c|}{\textbf{MinCut}} & \multicolumn{2}{|c|}{\textbf{FriendCut}} & \multicolumn{2}{|c|}{\textbf{EnemyCut}}\\
+\hline
+Patrol 1 & Patrol 2 & Patrol 1 & Patrol 2 & Patrol 1 & Patrol 2\\
+\hline
+Brandon & Christian & Brandon & Daniel & Brandon & Christian\\
+Cameron & Daniel & Cameron & Darwin & Cameron & Jake\\
+Colby & Jake & Christian & Jake & Colby & Jordan\\
+Darwin & Jordan & Colby & Jordan  & Daniel & Nathan\\
+Evan & Nathan & Evan & Nathan & Darwin & Patrick\\
+Tommy & Patrick & Tommy & Patrick & Evan & Timmy\\
+& Timmy & & Timmy & Tommy & \\
+\hline
+\multicolumn{6}{|c|}{\hphantom{1}} \\
+\hline
+\multicolumn{2}{|c|}{\textbf{AwkwardCut (+1)}} & \multicolumn{2}{|c|}{\textbf{AwkwardCut (+2)}} & \multicolumn{2}{|c|}{\textbf{HybridCut}} \\
+\hline
+Patrol 1 & Patrol 2 & Patrol 1 & Patrol 2 & Patrol 1 & Patrol 2\\
+\hline
+Brandon & Christian & Brandon & Cameron & Brandon & Christian\\
+Cameron & Daniel & Christian & Evan & Cameron & Daniel\\
+Colby & Jake & Daniel & Evan & Colby & Darwin\\
+Darwin & Jordan & Darwin & Nathan & Evan & Jake\\
+Evan & Nathan & Jake & Timmy & Timmy & Jordan\\
+Tommy & Patrick & Jordan & Tommy & Tommy & Nathan\\
+\hphantom{1} & Timmy & Patrick & \hphantom{1} & \hphantom{1} & Patrick\\
+\hline
+\end{tabular}
+\caption{The partitions arrived at by minimizing each objective function}
+\end{figure}
+
+For each method we define three scoring metrics:
+
+\begin{enumerate}
+\item \textbf{MinCut}.
+This is the traditional metric of a cut on a weighted graph and is the sum of the weights of the edges we would ``cut'' if we were to sever the edges between distinct patrols.
+More formally, for some partition $I, J$ $I \neq J$, the MinCut is defined as:
+$$
+MinCut(I, J) := \sum_{e = (i, j), i \in I, j \in J}^{} c(e)
+$$
+We want this to be as low as possible.
+\item \textbf{FriendCut}.
+This is the sum of the positive, or friendly edge weights within each patrol.
+That is, if $I_+ = \{e=(i,j) | i,j \in I, c(e)=1\}$ is the the set of edges in patrol $I$ with positive weight and $J_+ = \{e=(i,j) | i,j \in J, c(e)=1\}$ is the set of edges in patrol $J$ with positive weight, then (assuming all positive weights are 1):
+$$
+FriendCut(I, J) :=  |I_+ \cup J_+|
+$$
+We want this to be as high as possible.
+\item \textbf{EnemyCut}.
+This is similar to FriendCut, except we now measure the number of negative edges going from one scout to another within the same patrol.
+Let $I_- = \{e=(i,j) | i,j \in I, c(e)=-1\}$ be the the set of edges in patrol $I$ with negative weight and $J_- = \{e=(i,j) | i,j \in J, c(e)=-1\}$ be the set of edges in patrol $J$ with negative weight, then
+$$
+EnemyCut(I, J) := |I_- \cup J_-|
+$$
+    We want this to be as low as possible.
+\end{enumerate}
+
+The other methods are modifications of these three and are defined in Appendix A.
+
+\clearpage
+
+\begin{figure}[h]
+    \centering
+    \begin{tabular}{ |l|l|l|l| }
+        \hline
+        \textbf{Method} & \textbf{MinCut} & \textbf{FriendCut} & \textbf{EnemyCut} \\
+        \hline
+        MinCut & 0 & 17 & 0 \\
+        FriendCut & 1 & 18 & 1 \\
+        EnemyCut & 2 & 15 & 0 \\
+        AwkwardCut (+1) & 0 & 17 & 0 \\
+        AwkwardCut (+2) & 1 & 18 & 2 \\
+        HybridCut & 0 & 18 & 1 \\
+        \hline
+    \end{tabular}
+    \caption{The scores according to the three primary metrics.}
+\end{figure}
+
+Of the objective functions we chose to use, it seems like MinCut and HybridCut are the best performing.
+FriendCut performs almost as well as HybridCut, except for one worse in the EnemyCut objective.
+EnemyCut does poorly since finding a partition with 0 intra-patrol negative weight edges is easy in this example, and so EnemyCut settles for the first suboptimal (in terms of the other objective functions) solution it finds.
+AwkwardCut (+1) arrived at the same partition as MinCut (not altogether surprising, but informative, since even with the additional +1 penalty for placing Brandon and Tommy in the same patrol it decided it would be best off to do so anyways).
+But changing the ``awkward'' penalty to +2, the same penalty as mutual dislike in EnemyCut, we get a completely different partition.
+
+Let's take a more critical look at the optimal partition arrived at via MinCut:
+
+\begin{table}[h]
 	\caption{}\label{eqtable}
@@
 \end{table}
-\includegraphics[scale=0.65]{results1.JPG}
-Of the undesirable outcomes of such a partition, we can spot a few:
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.60]{graph}
+\end{figure}
+
 \begin{enumerate}
@@
 \end{enumerate}
-On the whole, though, there are no obvious improvements that can be made to our algorithm’s partition.
+
+On the whole, though, there are no obvious improvements that can be made to our algorithm's partition.
 We find that this is an assignment that could have reasonably been arrived at by Scoutmaster Bruce (granted the group dynamics have not changed since the scouts were surveyed).
-Furthermore, the solution took only 20 seconds to arrive at on a single core machine.
+Furthermore, the solution took only 10 seconds to arrive at on a single core machine.
 
@@
 There are $150!
-/ (10!)^{15}$ possible partitions, a number proportional to $10^164$ - meaning we could instead use that computational time to enumerate the number of atoms in the universe… a tredecillion $(10^{78})$ times.
+/ (10!)^{15}$ possible partitions, a number proportional to $10^{164}$ - meaning we could instead use that computational time to enumerate the number of atoms in the universe… a tredecillion $(10^{78})$ times.
 Knowing this, perhaps it won’t surprise the reader that this particular graph partitioning problem in general graphs is NP-complete - though there do exist approximation algorithms.%
@@
 
+\section*{Appendix A}
+The other objective functions.
+
+\begin{enumerate}
+    \item \textbf{AwkwardCut (+i)}.
+Same in all respects to MinCut, except for an additional penalty of $i$ if there are two scouts, Scout A and Scout B, in the same patrol such that Scout A likes Scout B and Scout B dislikes Scout A.
+    \begin{equation*}
+        \begin{split}
+            AwkwardCut&(I, J, i) := \Bigg[ \sum_{e = (k, j), k \in I, j \in J}^{} c(e) \Bigg] \\
+            &+ i * \left\vert{\{e_+ = (k,j), e_- = (j, k) : e_+, e_- \in I, c(e_+) = 1, c(e_-) = -1\}}\right\vert
+        \end{split}
+    \end{equation*}
+
+\item \textbf{HybridCut}. A convex combination of MinCut and FriendCut, each with equal weight (that is, half the MinCut loss plus half the FriendCut loss).
+    \begin{equation*}
+        HybridCut(I, J) := \frac{1}{2}MinCut(I, J) + \frac{1}{2}FriendCut(I, J)
+    \end{equation*}
+\end{enumerate}
+
 \bibliographystyle{amsplain}

### 2016-05-25 draft1.tex

@@
 %------------------------------------------------------------------------------
+=======
+
+\documentclass{amsart}
+
+%     If your article includes graphics, uncomment this command.
+\usepackage{graphicx}
+\usepackage{amsmath,amsfonts,amssymb,amscd,amsthm,amsbsy,upref}
+\usepackage[all]{xy}
+\usepackage{amsmath}
+\usepackage{mathrsfs}
+\usepackage{paralist}
+\usepackage{setspace}
+\usepackage{graphicx}
+\usepackage{tikz-cd}
+\usepackage{pgfplots}
+\usepackage{fancyhdr}
+\usepackage{tikz}
+\usepackage{pgfplots}
+\usepackage{array}
+
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}[theorem]{Lemma}
+
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{example}[theorem]{Example}
+\newtheorem{xca}[theorem]{Exercise}
+
+\theoremstyle{remark}
+\newtheorem{remark}[theorem]{Remark}
+
+\numberwithin{equation}{section}
+
+%    Absolute value notation
+\newcommand{\abs}[1]{\lvert#1\rvert}
+
+%    Blank box placeholder for figures (to avoid requiring any
+%    particular graphics capabilities for printing this document).
+\newcommand{\blankbox}[2]{%
+  \parbox{\columnwidth}{\centering
+%    Set fboxsep to 0 so that the actual size of the box will match the
+%    given measurements more closely.
+    \setlength{\fboxsep}{0pt}%
+    \fbox{\raisebox{0pt}[#2]{\hspace{#1}}}%
+  }%
+}
+
+\pgfplotsset{compat=1.13}
+
+\begin{document}
+
+\title{Assigning Scouts to Optimal Patrols}
+\author{Phil Snyder}
+\author{Ian Bruce}
+\author{Julio Marco Pineda}
+
+\begin{abstract}
+The Scoutmaster wants to find an arrangement of 13 new scouts into two patrols.
+The cohesiveness and overall quality experience of a patrol is severely affected by adverse relationship and can be improved by existing preferences.
+Thus, by using anonymous data from scouts and parents, a brute force strategy was employed to find the optimal arrangement of scouts that avoids severe conflict and maximizes existing preferences.
+An arrangement of 6 and 7 scouts on one patrol respectively was found that satisfies the desired conditions.
+\end{abstract}
+\maketitle
+\section*{Problem Description}
+Every year the Scoutmaster must assign new scouts to groups called patrols.
+In these patrols, scouts would plan and perform different scouting activities together in which they can form close friends that grow together.
+However, in recent years, the number of scouts have been increasing at a rate such that it is no longer logistically viable to place every new scout in a single patrol; multiple patrols are needed to organize activities better and to form proper bonds and community between the new recruits.
+Furthermore, the new scouts are in the fifth or sixth grade who have not fully developed their interpersonal skills.
+Many can be temperamental and awkward when dealing with adversity, and even some devastating conflicts occur where not only the two fighting scouts create animosity between each other, the rest of the troop alienates the destructive pair.
+If enough of these adverse interactions occur within a troop, many scouts would decide to quit and lose out in a great experience and community in the Boy Scouts.
+
+The data the scoutmaster collects about adverse pairings and preferences are either privately given by the scouts themselves or by their parents.
+The possibly destructive interactions are given by the parents of the scouts.Previously, the Scoutmaster was able to determine the perfect patrols manually due to having smaller number of scouts.
+However, the number of scouts have been increasing yearly to the point that solving this grouping problem cannot be done by hand in a reasonable time.
+Therefore, we can help our community partner find a better means to solve his troop organization problem.
+Our goal for this project is to assign boy scouts in different patrols of appropriate size (of 6-8 scouts) such that we maximize the retention rate of the scouts in the program by avoiding severe conflicts while maximizing positive relationships between scouts.
+
+We would like to answer the direct problem our community partner provided of arranging 15 scouts into two troops that avoids severe conflicts and promotes the positive relationships between the scouts.
+Furthermore, we want to determine if we can build a model to predict these patrol arrangements for any number of scouts so that the Scoutmaster can use this model for future years.
+Other than improving the overall experience of the scouts and alleviating the burden of the Scoutmaster, this model can possibly be applied to other situations where arranging a number of people of groups where interpersonal dynamics is necessary and important.
+
+\section*{Simplifications}
+Ultimately, our goal is to find groups for the scouts that will allow them to grow and be happy in the troop.
+If we really wanted to fully understand how a group can foster growth in an individual, we would want to consider the astronomical number of aspects about the personalities of each member and how those influence each of the other members given those other members’ astronomically complex personalities.
+Sociological and psychological literature would need to be studied extensively to learn about these aspects, and experts in the field would most likely be required to collect relevant data for each child, taking several weeks to conduct interviews.
+At the of the day, given our limited knowledge in this area, we settled with considering a subset of the sociological data between the scouts.
+We asked each scout which other scouts they liked and which ones they disliked.
+Now, this is obviously a giant simplification, but the difficulty of gathering more complex information and compiling said information would be too great for the expertise level of anyone involved in solving this problem.
+Also, this process of gathering data would have to be repeated each year for each new batch of new scouts, and would therefore be a burden on the Scoutmaster to collect a lot of data.
+
+Another simplification is to assume that the quality of a group can be determined by the quality of the relationships between the unique pairs of the group - in other words, the whole is equal to the sum of the parts.
+For example, consider the case where scouts A, B and C individually like each other when they are alone with one other scout, but don’t like being in a group together.
+This epiphenomenon won’t be considered; in context of our model, the pairing of these scouts would be heavily favored.
+
+\section*{Mathematical Model}
+This problem visually lends itself well to the notion of a mathematical graph.
+Since there is an inherent sense of relationships between individuals (how much a scout likes another scout), it is appropriate to consider the scouts as vertices in a graph with directed edges representing how much the “head” scout likes the “tail” scout.
+Our goal, using this language, is to find a partition of the vertices in the graph that satisfies some sort of objective function given size constraints of the partitions.
+In our problem at hand, we consider a complete graph with 13 vertices and edge weights being either 0, 1, or -1, corresponding to indifference, approval, or disapproval respectively of the “head” scout with respect to the “tail” scout.
+The Scoutmaster specifically wants two groups out of these 13, and with the size constraints for a patrol considered, our goal becomes finding 6 vertices to be in one partition and the other 7 being in another.
+The total number of partitions becomes 13 choose 6, which is equivalent to 1,716.
+This is certainly within the scope of considering total enumeration of the partitions as a reasonable solution to the problem, therefore we will proceed accordingly and implement the calculation using Python and its associated libraries to represent the data in a format that is convenient to work with.
+
+Now, a natural question for us to consider is what objective function would be most appropriate to optimize.
+A natural answer to this would be to sum the edge weights of all edges that connect vertices that are within partitions.
+This would make sense if separating enemies is just as important as joining friends, but this just might not be the case.
+In terms of weighing these factors, we defer to Scoutmaster Bruce’s experience with how past patrols fared with friends and foes in the mix.
+Therefore, we consider a family of possible objective functions, namely giving real numbered weights to instances of approving relationships and disapproving relationships, and select a few of these objective functions to solve for.
+We then present the collection of results from different objective functions to the Scoutmaster for him to see which one seems to most reasonably match his preferences.
+\section*{Solution of Mathematical Problem}
+We represent our directed graph as a 13 by 13 similarity matrix, or, as might be more fitting for the problem, an “affinity” matrix. Like below:%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\footnote{The scouts were enumerated like so: 1. Brandon, 2. Cameron, 3, Christian, 4. Colby, 5. Daniel, 6. Darwin, 7. Evan, 8. Jake, 9. Jordan, 10. Nathan, 11. Patrick, 12. Timmy, 13. Tommy.}%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\includegraphics[scale=0.6]{solution1.JPG}
+
+The NA value represents a hard constraint established by a scout’s parent (i.e., under no circumstances are these two to be in the same patrol).
+In our implementation we convert the NA value to $-\infty$.
+We designed an objective function that accepts two scouts who are assumed to be in different patrols and calculates the loss associated with such a placement.
+The loss is defined to be $L(x, y) = \text{max}(-2, x[y] + y[x])$ where $x[y]$ is $x$’s disposition towards $y$ and vise versa.
+For example, if $x$ dislikes $y$ and $y$ dislikes $x$, the loss is -2.
+If $x$ likes $y$ but $y$ dislikes $x$ the loss is 0.
+If $x$ and $y$ both like each other, the loss is 2.
+In mixed cases where the first scout is indifferent to another, but the other likes or dislikes the first scout, the loss is +1 or -1 respectively.
+We are trying to minimize the loss, so in applying our loss function over the entire dataset, our algorithm is equally averse to same-patrol mutual enmity as it is to dividing mutual friendship, half as averse to the indifference of one scout and the stronger inclination of another, and indifferent to polar opposite affinities within the same patrol.
+There is a question as to whether this is an accurate reflection of how individual scouts would view their own “loss” of any of the possible scenarios.
+As it turns out, our algorithm arrives at a very reasonable solution regardless.
+\section*{Results}
+As mentioned earlier, the number of scouts we must partition is small enough to find an optimal solution by brute force.
+Our algorithm finds a partition of scouts into patrols of sizes 6 and 7 with a total loss of 0 as described in the table below:
+\begin{table}[ht]
+	\caption{}\label{eqtable}
+	\renewcommand\arraystretch{1.5}
+	\noindent\[
+	\begin{array}{|c|c|}
+	\hline
+	\textbf{Patrol 1}&\textbf{Patrol 2}\\
+	\hline
+	\text{Brandon, Cameron, Colby,}&\text{Christian, Daniel, Jake,}\\
+	\text{Darwin, Evan, Tommy}&\text{Jordan, Nathan, Patrick, Timmy}\\
+	\hline
+	\end{array}
+	\]
+\end{table}
+\includegraphics[scale=0.65]{results1.JPG}
+Of the undesirable outcomes of such a partition, we can spot a few:
+\begin{enumerate}
+	\item Tommy and Timmy both like each other and are in separate patrols.
+But if we move Timmy to Patrol 1 ($P1$), we break up the mutual friendship of Timmy and Nathan in $P2$.
+Suppose we moved both Timmy and Nathan to $P1$ and Brandon (who dislikes Nathan) to $P2$.
+But then Brandon loses two friends (Tommy and Cameron), gains one (Jordan), and now neither Nathan nor Timmy are in the same patrol as Daniel, whom they both like.
+Although, Brandon and Tommy find themselves in the awkward situation of having complete opposite sentiments towards each other.
+Perhaps this would be a worthwhile trade after all (and a larger, negative weight needs to be placed upon disfavor-favor relationships).
+
+	\item As mentioned in the previous point, Tommy dislikes Brandon.
+Trading Tommy in $P1$ for Jordan in $P2$ seems promising, But we have overlooked the fact that Colby and Jordan are our NA pair, and cannot be in the same patrol under any circumstances.
+Trading Colby for Jordan is another option, but Colby will be missed by Darwin, Evan, Cameron and Tommy (In other words, the whole of P1 minus Brandon).
+\end{enumerate}
+On the whole, though, there are no obvious improvements that can be made to our algorithm’s partition.
+We find that this is an assignment that could have reasonably been arrived at by Scoutmaster Bruce (granted the group dynamics have not changed since the scouts were surveyed).
+Furthermore, the solution took only 20 seconds to arrive at on a single core machine.
+
+\section*{Improvements}
+While our algorithm demonstrably works well on smaller groups of 13 people, we do not expect our brute force solution to be tractable on larger groups wherein greater than 2 partitions are required.
+As a possible scenario, consider a company of 150 people to be divided into fixed-size teams of 10 individuals each.
+There are $150!
+/ (10!)^{15}$ possible partitions, a number proportional to $10^164$ - meaning we could instead use that computational time to enumerate the number of atoms in the universe… a tredecillion $(10^{78})$ times.
+Knowing this, perhaps it won’t surprise the reader that this particular graph partitioning problem in general graphs is NP-complete - though there do exist approximation algorithms.%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\footnote{B.
+W.
+Kernighan, S.
+Lin, \textit{An Efficient Heuristic Procedure for Partitioning Graphs}.
+Bell System Technical Journal.
+\textbf{49} (1970), 291--307.
+doi: 10.1002/j.1538-7305.1970.tb01770.x}%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+As scout enrollment increases, it becomes necessary to consider scenarios where we have both a large number of scouts in need of assignment and multiple patrols to choose from.
+Our algorithm could be extended to use brute force to find a globally optimal solution when it is deemed computationally feasible and to use an approximating algorithm to find a locally optimal solution otherwise.
+
+\section*{Conclusions}
+We initially struggled to decide on determining an appropriate objective function to partition the scouts to different groups.
+Follow up interviews with Scoutmaster Gene Bruce allowed us to decide on this function.
+We learned to rely on his expertise and intuition rather than trying to blindly decide on which types of relationships should be fostered and avoided to weight our objective function.
+Thus, we learned necessary communication skills and consideration of our community partner's needs.
+
+With a feasibility check of how many solutions possible once the mathematical model was decided using graphs, a brute force strategy was the most accessible method to arrive at a solution.
+Thus we were able to develop a program that can partition scouts using this strategy.
+We learned to utilize tools we have at hand right away, and then we can consider other methods later on to reduce the computational time of our algorithm.
+
+\section*{Acknowledgments}
+We would like to thank Scoutmaster Gene Bruce for providing us the necessary data and being extremely patient and understanding when discussing the background and problem.
+We would also like to thank Professor Sarah Billey for the mentorship and guidance she provided throughout this project, and our TA Austin Tran for his comments and suggestions.
+
+\section*{Verification Statement}
+(Once we receive our verification statement, we will place it here.)
+
+\bibliographystyle{amsplain}
+\begin{thebibliography}{10}
+
+\bibitem {A} B.
+W.
+Kernighan, S.
+Lin, \textit{An Efficient Heuristic Procedure for Partitioning Graphs}.
+Bell System Technical Journal.
+\textbf{49} (1970), 291--307.
+doi: 10.1002/j.1538-7305.1970.tb01770.x
+
+\end{thebibliography}
+
+\end{document}
+
+%------------------------------------------------------------------------------
+% End of journal.tex
+%------------------------------------------------------------------------------
+>>>>>>> 9771af696fbc57423120850744607162fcf171d5

### 2016-05-25 approximate.py
Starter code for approximation

@@ -1,6 +1,55 @@
 import numpy as np
+import itertools
 
 def rand_affinity(n, probs):
 	return np.random.choice([-1,0,1], (n,n), p=probs)
 
-print(rand_affinity(10, [.1, .7, .2]))
+mat = rand_affinity(200, [.1, .7, .2])
+np.savetxt("test_mat.csv", mat, delimiter=",")
+
+# In this algorithm, we will represent partitions as incidence vectors, with 1
+# at position i representing scout i being in the partition, and 0 representing
+# them not being in the partition
+def compute_sol(matrix):
+
+	# This function expects 1's at incidences of scouts being in the first partition,
+	# 2's where scouts are in the second, and 0's otherwise
+	def optimize_pair(part):
+		G = 1
+		Ds = [[diff(i, part) for i in range(len(part))] for j in range(1,3)] # Computes D's
+		return Ds
+
+	# i - the index of the scout in question
+	# part - the partition scout i is a member of
+	def external(i, part):
+		sum = 0
+		for j in range(len(part)):
+			sum += matrix[i, j] if part[j] == 2 else 0
+		return sum
+
+	# i - the index of the scout in question
+	# part - the partition scout i is a member of
+	def internal(i, part):
+		sum = 0
+		for j in range(len(part)):
+			sum += matrix[i, j] if part[j] == 1 else 0
+		return sum
+
+	# difference between external and internal costs for the
+	def diff(i, part):
+		return external(i, part) - internal(i, part)
+
+	def gain(d1, d2, i, j):
+		return d1 + d2 - matrix[i,j]
+
+	def update_ext():
+		pass
+
+	def swap_cost(a, b):
+		pass
+
+
+	return optimize_pair([i for i in itertools.chain(itertools.repeat(1, 100), itertools.repeat(2, 100))])
+
+print(compute_sol(mat))
+

### 2016-05-25 draft1.tex
Updated Abstract and Problem Description

@@
 \begin{abstract}
-The Scoutmaster wants to find an arrangement of 13 new scouts into two patrols.
-The cohesiveness and overall quality experience of a patrol is severely affected by adverse relationship and can be improved by existing preferences.
+Scoutmaster Gene Bruce of the Boy Scouts of America for Troop 407 in Kent, Washington wants to find an arrangement of 13 new scouts into two patrols.
+The cohesiveness and overall quality experience of a patrol is affected by adverse relationships and existing friendships.
 Thus, by using anonymous data from scouts and parents, a brute force strategy was employed to find the optimal arrangement of scouts that avoids severe conflict and maximizes existing preferences.
-An arrangement of 6 and 7 scouts on one patrol respectively was found that satisfies the desired conditions.
+We found an optimal arrangement of the 13 boy scouts that satisfies the desired conditions.
 \end{abstract}
@@
 
-The data the scoutmaster collects about adverse pairings and preferences are either privately given by the scouts themselves or by their parents.
-The possibly destructive interactions are given by the parents of the scouts.Previously, the Scoutmaster was able to determine the perfect patrols manually due to having smaller number of scouts.
-However, the number of scouts have been increasing yearly to the point that solving this grouping problem cannot be done by hand in a reasonable time.
+The data the Scoutmaster collects about adverse pairings and preferences are either privately given by the scouts themselves.
+Possibly destructive interactions are given by the parents.
+In the past, the Scoutmaster determined the perfect arrangement of patrols without too much time and effort due to the smaller number of new recruits.
+However, newer scouts are joining yearly to the point that solving this grouping problem cannot be done by hand in a reasonable time.
 Therefore, we can help our community partner find a better means to solve his troop organization problem.
@@
 
-We would like to answer the direct problem our community partner provided of arranging 15 scouts into two troops that avoids severe conflicts and promotes the positive relationships between the scouts.
+We aim to answer the direct problem our community partner provided of arranging 13 scouts into two troops that avoids severe conflicts and promotes the positive relationships between the scouts.
 Furthermore, we want to determine if we can build a model to predict these patrol arrangements for any number of scouts so that the Scoutmaster can use this model for future years.
-Other than improving the overall experience of the scouts and alleviating the burden of the Scoutmaster, this model can possibly be applied to other situations where arranging a number of people of groups where interpersonal dynamics is necessary and important.
+Additionally, this model can possibly be applied to other situations in which arranging a number of people into groups where interpersonal dynamics is necessary and important such as project teams in a company and forming new military troops.
 

### 2016-05-25 draft1.tex
More edits
Added "Data" section and "Appendix B" talking about Stirling number.
Updated Conclusion section.

@@
 
-\pgfplotsset{compat=1.13}
-
 \begin{document}
@@
 Additionally, this model can possibly be applied to other situations in which arranging a number of people into groups where interpersonal dynamics is necessary and important such as project teams in a company and forming new military troops.
+
+\section*{Data}
+The 13 different recruited scouts are Brandon, Cameron, Christian, Colby, Daniel, Darwin, Evan, Jake, Jordan, Nathan, Patrick, Timmy and Tommy.
+After the Scoutmaster collected the likes and dislikes of each scout, he presented the data as follows:
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.60]{graph}
+\end{figure}
+
+The solid arrow represents a preferred companion.
+For example, there is a solid arrow from Tommy pointing to Colby indicating that Tommy likes to be with Colby.
+A dashed arrow represents a dislike.
+For example, there is a dashed arrow from Tommy to Jordan showing that Tommy does not want to be with Jordan.
+Furthermore, the Scoutmaster provided us a separate email, and indicated that Colby and Jordan cannot be in the same troop.
+The highlighted arrows in the figure above does not mean anything new between the relationships of the boy scouts.
+
+The Stirling number of the second kind provides the total possible ways to arrange any number of distinguishable objects to indistinguishable groups as long as any troop size is permitted and no possible empty troops.
+For the data in this project, the Stirling number of the second kind for 13 scouts into 2 troops is 4095.
+This number provides an upper bound on the possible arrangements for the specific problem tackled in this project.
 
@@
 \end{table}
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.60]{graph}
-\end{figure}
 
@@
 \end{enumerate}
+
+\section*{Appendix B}
+The solution for finding the Stirling number of the second kinds for $n = 13$ scouts to be arranged in $k = 2$ troops:
+\begin{gather*}
+	S(n,k) = \frac{1}{k!} \sum_{i=0}^{k}(-1)^k {2 \choose i} (k - i)^{n} \\
+	S(13,2) = \frac{1}{2!} \sum_{i=0}^{2}(-1)^2 {2 \choose i} (2 - i)^{13} \\
+	S(13, 2) = 4095
+\end{gather*}
 
@@
 \section*{Conclusions}
-We initially struggled to decide on determining an appropriate objective function to partition the scouts to different groups.
+We initially struggled deciding on an appropriate objective function to partition the scouts to different groups.
 Follow up interviews with Scoutmaster Gene Bruce allowed us to decide on this function.
-We learned to rely on his expertise and intuition rather than trying to blindly decide on which types of relationships should be fostered and avoided to weight our objective function.
+We relied on his expertise and intuition rather than trying to blindly decide on which types of relationships should be fostered and avoided to determine our objective function.
 Thus, we learned necessary communication skills and consideration of our community partner's needs.
 
-With a feasibility check of how many solutions possible once the mathematical model was decided using graphs, a brute force strategy was the most accessible method to arrive at a solution.
+With a feasibility check of how many solutions possible once the mathematical model was decided using graphs, a brute force strategy was the most reliable method to arrive at a solution.
 Thus we were able to develop a program that can partition scouts using this strategy.
-We learned to utilize tools we have at hand right away, and then we can consider other methods later on to reduce the computational time of our algorithm.
+We learned to utilize tools we have at hand.
+Further consideration of other methods later on to reduce the computational time of our algorithm.
 
@@
 We would like to thank Scoutmaster Gene Bruce for providing us the necessary data and being extremely patient and understanding when discussing the background and problem.
-We would also like to thank Professor Sarah Billey for the mentorship and guidance she provided throughout this project, and our TA Austin Tran for his comments and suggestions.
+We would also like to thank Professor Sara Billey for the mentorship and guidance she provided throughout this project, and our TA Austin Tran for his comments and suggestions.
 

### 2016-05-25 draft1.tex
Adding Previous of Work section
Still have to add actual text. Only outline is there.

@@
 This number provides an upper bound on the possible arrangements for the specific problem tackled in this project.
+
+\section*{Previous Work}
+(Currently reading more on Max Flow-Min cut, K-means clustering, Graph clustering: http://www.sciencedirect.com/science/article/pii/S1574013707000020. We will talk about generating random solutions)
 
@@
 \section*{Appendix B}
-The solution for finding the Stirling number of the second kinds for $n = 13$ scouts to be arranged in $k = 2$ troops:
+The solution for finding the Stirling number of the second kind for $n = 13$ scouts to be arranged in $k = 2$ troops:
 \begin{gather*}

### 2016-05-25 draft1.tex
Quick edits on Problem Description

@@
 \section*{Problem Description}
-Every year the Scoutmaster must assign new scouts to groups called patrols.
-In these patrols, scouts would plan and perform different scouting activities together in which they can form close friends that grow together.
-However, in recent years, the number of scouts have been increasing at a rate such that it is no longer logistically viable to place every new scout in a single patrol; multiple patrols are needed to organize activities better and to form proper bonds and community between the new recruits.
-Furthermore, the new scouts are in the fifth or sixth grade who have not fully developed their interpersonal skills.
+The community partner we are serving for this project is Scoutmaster Gene Bruce of the Boys Scouts of America.
+He leads Troop 407 in Kent, Washington and every year he must assign new scouts to groups called patrols.
+In these patrols, scouts would plan, perform different scouting activities together and form close bonds.
+However, in recent years, the number of scouts have been increasing at such a rate that it is no longer logistically viable to place every new scout in a single patrol; multiple patrols are needed to organize activities better and to form a community between the new recruits.
+Furthermore, the new scouts are in the fifth or sixth grade and have not fully developed their interpersonal skills.
 Many can be temperamental and awkward when dealing with adversity, and even some devastating conflicts occur where not only the two fighting scouts create animosity between each other, the rest of the troop alienates the destructive pair.

### 2016-05-25 draft1.tex
changes to paper

--- 2016-05-25 draft1.tex
@@
 
+\pgfplotsset{compat=1.13}
+
 \begin{document}
@@
 \begin{abstract}
-Scoutmaster Gene Bruce of the Boy Scouts of America for Troop 407 in Kent, Washington wants to find an arrangement of 13 new scouts into two patrols.
-The cohesiveness and overall quality experience of a patrol is affected by adverse relationships and existing friendships.
+The Scoutmaster wants to find an arrangement of 13 new scouts into two patrols.
+The cohesiveness and overall quality experience of a patrol is severely affected by adverse relationship and can be improved by existing preferences.
 Thus, by using anonymous data from scouts and parents, a brute force strategy was employed to find the optimal arrangement of scouts that avoids severe conflict and maximizes existing preferences.
-We found an optimal arrangement of the 13 boy scouts that satisfies the desired conditions.
+An arrangement of 6 and 7 scouts on one patrol respectively was found that satisfies the desired conditions.
 \end{abstract}
@@
 
-The data the Scoutmaster collects about adverse pairings and preferences are either privately given by the scouts themselves.
-Possibly destructive interactions are given by the parents.
-In the past, the Scoutmaster determined the perfect arrangement of patrols without too much time and effort due to the smaller number of new recruits.
-However, newer scouts are joining yearly to the point that solving this grouping problem cannot be done by hand in a reasonable time.
+The data the scoutmaster collects about adverse pairings and preferences are either privately given by the scouts themselves or by their parents.
+The possibly destructive interactions are given by the parents of the scouts.Previously, the Scoutmaster was able to determine the perfect patrols manually due to having smaller number of scouts.
+However, the number of scouts have been increasing yearly to the point that solving this grouping problem cannot be done by hand in a reasonable time.
 Therefore, we can help our community partner find a better means to solve his troop organization problem.
@@
 
-We aim to answer the direct problem our community partner provided of arranging 13 scouts into two troops that avoids severe conflicts and promotes the positive relationships between the scouts.
+We would like to answer the direct problem our community partner provided of arranging 15 scouts into two troops that avoids severe conflicts and promotes the positive relationships between the scouts.
 Furthermore, we want to determine if we can build a model to predict these patrol arrangements for any number of scouts so that the Scoutmaster can use this model for future years.
-Additionally, this model can possibly be applied to other situations in which arranging a number of people into groups where interpersonal dynamics is necessary and important such as project teams in a company and forming new military troops.
-
-\section*{Data}
-The 13 different recruited scouts are Brandon, Cameron, Christian, Colby, Daniel, Darwin, Evan, Jake, Jordan, Nathan, Patrick, Timmy and Tommy.
-After the Scoutmaster collected the likes and dislikes of each scout, he presented the data as follows:
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=0.60]{graph}
-\end{figure}
-
-The solid arrow represents a preferred companion.
-For example, there is a solid arrow from Tommy pointing to Colby indicating that Tommy likes to be with Colby.
-A dashed arrow represents a dislike.
-For example, there is a dashed arrow from Tommy to Jordan showing that Tommy does not want to be with Jordan.
-Furthermore, the Scoutmaster provided us a separate email, and indicated that Colby and Jordan cannot be in the same troop.
-The highlighted arrows in the figure above does not mean anything new between the relationships of the boy scouts.
-
-The Stirling number of the second kind provides the total possible ways to arrange any number of distinguishable objects to indistinguishable groups as long as any troop size is permitted and no possible empty troops.
-For the data in this project, the Stirling number of the second kind for 13 scouts into 2 troops is 4095.
-This number provides an upper bound on the possible arrangements for the specific problem tackled in this project.
+Other than improving the overall experience of the scouts and alleviating the burden of the Scoutmaster, this model can possibly be applied to other situations where arranging a number of people of groups where interpersonal dynamics is necessary and important.
 
@@
 \end{table}
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.60]{graph}
+\end{figure}
 
@@
 \end{enumerate}
-
-\section*{Appendix B}
-The solution for finding the Stirling number of the second kinds for $n = 13$ scouts to be arranged in $k = 2$ troops:
-\begin{gather*}
-	S(n,k) = \frac{1}{k!} \sum_{i=0}^{k}(-1)^k {2 \choose i} (k - i)^{n} \\
-	S(13,2) = \frac{1}{2!} \sum_{i=0}^{2}(-1)^2 {2 \choose i} (2 - i)^{13} \\
-	S(13, 2) = 4095
-\end{gather*}
 
@@
 \section*{Mathematical Model}
-This problem visually lends itself well to the notion of a mathematical graph.
+The problem of placing individuals into groups visually lends itself well to the notion of a mathematical graph.
+In mathematical vocabulary, a graph consists of a group of objects called vertices that get connected with an "edge" if they share some sort of relationship.
+The graph we will use to represent the data of our problem can be visualized with the image offered on page (5?), where vertices are represented by the circled names and edges are represented by lines between the circles.
 Since there is an inherent sense of relationships between individuals (how much a scout likes another scout), it is appropriate to consider the scouts as vertices in a graph with directed edges representing how much the “head” scout likes the “tail” scout.
@@
 \section*{Conclusions}
-We initially struggled deciding on an appropriate objective function to partition the scouts to different groups.
+We initially struggled to decide on determining an appropriate objective function to partition the scouts to different groups.
 Follow up interviews with Scoutmaster Gene Bruce allowed us to decide on this function.
-We relied on his expertise and intuition rather than trying to blindly decide on which types of relationships should be fostered and avoided to determine our objective function.
+We learned to rely on his expertise and intuition rather than trying to blindly decide on which types of relationships should be fostered and avoided to weight our objective function.
 Thus, we learned necessary communication skills and consideration of our community partner's needs.
 
-With a feasibility check of how many solutions possible once the mathematical model was decided using graphs, a brute force strategy was the most reliable method to arrive at a solution.
+With a feasibility check of how many solutions possible once the mathematical model was decided using graphs, a brute force strategy was the most accessible method to arrive at a solution.
 Thus we were able to develop a program that can partition scouts using this strategy.
-We learned to utilize tools we have at hand.
-Further consideration of other methods later on to reduce the computational time of our algorithm.
+We learned to utilize tools we have at hand right away, and then we can consider other methods later on to reduce the computational time of our algorithm.
 
@@
 We would like to thank Scoutmaster Gene Bruce for providing us the necessary data and being extremely patient and understanding when discussing the background and problem.
-We would also like to thank Professor Sara Billey for the mentorship and guidance she provided throughout this project, and our TA Austin Tran for his comments and suggestions.
+We would also like to thank Professor Sarah Billey for the mentorship and guidance she provided throughout this project, and our TA Austin Tran for his comments and suggestions.
 

### 2016-05-25 approximate.py

@@ -1,5 +1,6 @@
 import numpy as np
 import itertools
+import pandas as pd
 
 def rand_affinity(n, probs):
 	return np.random.choice([-1,0,1], (n,n), p=probs)
@@ -17,7 +18,8 @@ def compute_sol(matrix):
 	def optimize_pair(part):
 		G = 1
 		Ds = [[diff(i, part) for i in range(len(part))] for j in range(1,3)] # Computes D's
-		return Ds
+		
+
 
 	# i - the index of the scout in question
 	# part - the partition scout i is a member of

### 2016-05-25 draft1.tex

--- 2016-05-25 draft1.tex
@@
+\documentclass{amsart}
+
+%     If your article includes graphics, uncomment this command.
+\usepackage{graphicx}
+\usepackage{amsmath,amsfonts,amssymb,amscd,amsthm,amsbsy,upref}
+\usepackage[all]{xy}
+\usepackage{amsmath}
+\usepackage{mathrsfs}
+\usepackage{paralist}
+\usepackage{setspace}
+\usepackage{graphicx}
+\usepackage{tikz-cd}
+\usepackage{pgfplots}
+\usepackage{fancyhdr}
+\usepackage{tikz}
+\usepackage{pgfplots}
+\usepackage{array}
+
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}[theorem]{Lemma}
+
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{example}[theorem]{Example}
+\newtheorem{xca}[theorem]{Exercise}
+
+\theoremstyle{remark}
+\newtheorem{remark}[theorem]{Remark}
+
+\numberwithin{equation}{section}
+
+%    Absolute value notation
+\newcommand{\abs}[1]{\lvert#1\rvert}
+
+%    Blank box placeholder for figures (to avoid requiring any
+%    particular graphics capabilities for printing this document).
+\newcommand{\blankbox}[2]{%
+  \parbox{\columnwidth}{\centering
+%    Set fboxsep to 0 so that the actual size of the box will match the
+%    given measurements more closely.
+    \setlength{\fboxsep}{0pt}%
+    \fbox{\raisebox{0pt}[#2]{\hspace{#1}}}%
+  }%
+}
+
+\pgfplotsset{compat=1.13}
+
+\begin{document}
+
+\title{Assigning Scouts to Optimal Patrols}
+\author{Phil Snyder}
+\author{Ian Bruce}
+\author{Julio Marco Pineda}
+
+\begin{abstract}
+The Scoutmaster wants to find an arrangement of 13 new scouts into two patrols.
+The cohesiveness and overall quality experience of a patrol is severely affected by adverse relationship and can be improved by existing preferences.
+Thus, by using anonymous data from scouts and parents, a brute force strategy was employed to find the optimal arrangement of scouts that avoids severe conflict and maximizes existing preferences.
+An arrangement of 6 and 7 scouts on one patrol respectively was found that satisfies the desired conditions.
+\end{abstract}
+\maketitle
+\section*{Problem Description}
+Every year the Scoutmaster must assign new scouts to groups called patrols.
+In these patrols, scouts would plan and perform different scouting activities together in which they can form close friends that grow together.
+However, in recent years, the number of scouts have been increasing at a rate such that it is no longer logistically viable to place every new scout in a single patrol; multiple patrols are needed to organize activities better and to form proper bonds and community between the new recruits.
+Furthermore, the new scouts are in the fifth or sixth grade who have not fully developed their interpersonal skills.
+Many can be temperamental and awkward when dealing with adversity, and even some devastating conflicts occur where not only the two fighting scouts create animosity between each other, the rest of the troop alienates the destructive pair.
+If enough of these adverse interactions occur within a troop, many scouts would decide to quit and lose out in a great experience and community in the Boy Scouts.
+
+The data the scoutmaster collects about adverse pairings and preferences are either privately given by the scouts themselves or by their parents.
+The possibly destructive interactions are given by the parents of the scouts.Previously, the Scoutmaster was able to determine the perfect patrols manually due to having smaller number of scouts.
+However, the number of scouts have been increasing yearly to the point that solving this grouping problem cannot be done by hand in a reasonable time.
+Therefore, we can help our community partner find a better means to solve his troop organization problem.
+Our goal for this project is to assign boy scouts in different patrols of appropriate size (of 6-8 scouts) such that we maximize the retention rate of the scouts in the program by avoiding severe conflicts while maximizing positive relationships between scouts.
+
+We would like to answer the direct problem our community partner provided of arranging 15 scouts into two troops that avoids severe conflicts and promotes the positive relationships between the scouts.
+Furthermore, we want to determine if we can build a model to predict these patrol arrangements for any number of scouts so that the Scoutmaster can use this model for future years.
+Other than improving the overall experience of the scouts and alleviating the burden of the Scoutmaster, this model can possibly be applied to other situations where arranging a number of people of groups where interpersonal dynamics is necessary and important.
+
+\section*{Simplifications}
+Ultimately, our goal is to find groups for the scouts that will allow them to grow and be happy in the troop.
+If we really wanted to fully understand how a group can foster growth in an individual, we would want to consider the astronomical number of aspects about the personalities of each member and how those influence each of the other members given those other members’ astronomically complex personalities.
+Sociological and psychological literature would need to be studied extensively to learn about these aspects, and experts in the field would most likely be required to collect relevant data for each child, taking several weeks to conduct interviews.
+At the of the day, given our limited knowledge in this area, we settled with considering a subset of the sociological data between the scouts.
+We asked each scout which other scouts they liked and which ones they disliked.
+Now, this is obviously a giant simplification, but the difficulty of gathering more complex information and compiling said information would be too great for the expertise level of anyone involved in solving this problem.
+Also, this process of gathering data would have to be repeated each year for each new batch of new scouts, and would therefore be a burden on the Scoutmaster to collect a lot of data.
+
+Another simplification is to assume that the quality of a group can be determined by the quality of the relationships between the unique pairs of the group - in other words, the whole is equal to the sum of the parts.
+For example, consider the case where scouts A, B and C individually like each other when they are alone with one other scout, but don’t like being in a group together.
+This epiphenomenon won’t be considered; in context of our model, the pairing of these scouts would be heavily favored.
+
+\section*{Mathematical Model}
+To model the relationships between scouts, we use a mathematical graph.
+We let the scouts be vertices in our graph.
+A directed edge between Scout A and Scout B has a weight proportional to how much Scout A likes Scout B.
+We would then like to find a partition of the vertices in the graph that minimizes an objective function given constraints on the sizes of subsets generated by our partition.
+This results in a complete graph on 13 vertices having edge weights 0, 1, or -1, corresponding to indifference, approval, or disapproval, respectively.
+Scoutmaster Bruce specifically would like two patrols from the 13 scouts - one of size 6, the other of size 7.
+The total number of partitions given these constrants is $\binom{13}{6} = \binom{13}{7} = 1716$.
+This number is reasonable enough that we may calculate the total loss of every valid grouping and select the partition with the lowest score.
+
+We measure the goodness (or score) of a partition in six different ways, corresponding to six different objective functions, which we dub MinCut, FriendCut, EnemyCut, AwkwardCut, and HybridCut.
+MinCut is designed to minimize the cut edge weights between partitions (see \textit{Solution of Mathematical Problem}).
+FriendCut is designed to maximize the number of intra-patrol positive edge weights.
+EnemyCut is meant to minimize the number of intra-patrol negative edge weights, Awkward cut ($+i$) is the same as MinCut, but with an additional penalty of $i$ if there are a pair of scouts, Scout A and Scout B, such that Scout A favors Scout B, but Scout B disfavors Scout A.
+Hybrid Cut is a convex combination of MinCut and FriendCut, each with equal weight (that is, half the MinCut loss plus half the FriendCut loss).
+Our three primary objective functions - MinCut, FriendCut, and EnemyCut are given a more rigorous treatment in the next section.
+
+\section*{Solution of Mathematical Problem}
+We represent our directed graph as a 13 by 13 similarity matrix, or, as might
+be more fitting for the problem, an ``affinity'' matrix.
+Like below\footnote{The scouts were enumerated like so: 1.
+Brandon, 2.
+Cameron, 3, Christian, 4.
+Colby, 5.
+Daniel, 6.
+Darwin, 7.
+Evan, 8.
+Jake, 9.
+Jordan, 10.
+Nathan, 11.
+Patrick, 12.
+Timmy, 13.
+Tommy.}:
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.28]{data}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The NA value represents a hard constraint established by a scout's parent (i.e., under no circumstances are these two to be in the same patrol).
+We designed a number of objective functions that accept two scouts and calculates the loss associated with such a placement.
+The loss is defined to be $L(x, y) = \text{max}(-2, x[y] + y[x])$ where $x[y]$ is $x$'s disposition towards $y$ and vise versa.
+For example, if $x$ dislikes $y$ and $y$ dislikes $x$, the loss is -2.
+If $x$ likes $y$ but $y$ dislikes $x$ the loss is 0.
+If $x$ and $y$ both like each other, the loss is 2.
+In mixed cases where the first scout is indifferent to another, but the other likes or dislikes the first scout, the loss is +1 or -1 respectively.
+We are trying to minimize the loss, so in applying our loss function over the entire dataset, our algorithm is equally averse to same-patrol mutual enmity as it is to dividing mutual friendship, half as averse to the indifference of one scout and the stronger inclination of another, and indifferent to polar opposite affinities within the same patrol.
+There is a question as to whether this is an accurate reflection of how individual scouts would view their own ``loss'' of any of the possible scenarios.
+As it turns out, our algorithm arrives at a very reasonable solution regardless.
+\section*{Results}
+As mentioned earlier, the number of scouts we must partition is small enough to find an optimal solution by brute force.
+Our algorithm finds a partition of scouts into patrols of sizes 6 and 7 which is optimal with respect to each objective.
+
+\clearpage
+
+\begin{figure}[h]
+\centering
+\begin{tabular}{ |l|l|l|l|l|l| }
+\hline
+\multicolumn{2}{|c|}{\textbf{MinCut}} & \multicolumn{2}{|c|}{\textbf{FriendCut}} & \multicolumn{2}{|c|}{\textbf{EnemyCut}}\\
+\hline
+Patrol 1 & Patrol 2 & Patrol 1 & Patrol 2 & Patrol 1 & Patrol 2\\
+\hline
+Brandon & Christian & Brandon & Daniel & Brandon & Christian\\
+Cameron & Daniel & Cameron & Darwin & Cameron & Jake\\
+Colby & Jake & Christian & Jake & Colby & Jordan\\
+Darwin & Jordan & Colby & Jordan  & Daniel & Nathan\\
+Evan & Nathan & Evan & Nathan & Darwin & Patrick\\
+Tommy & Patrick & Tommy & Patrick & Evan & Timmy\\
+& Timmy & & Timmy & Tommy & \\
+\hline
+\multicolumn{6}{|c|}{\hphantom{1}} \\
+\hline
+\multicolumn{2}{|c|}{\textbf{AwkwardCut (+1)}} & \multicolumn{2}{|c|}{\textbf{AwkwardCut (+2)}} & \multicolumn{2}{|c|}{\textbf{HybridCut}} \\
+\hline
+Patrol 1 & Patrol 2 & Patrol 1 & Patrol 2 & Patrol 1 & Patrol 2\\
+\hline
+Brandon & Christian & Brandon & Cameron & Brandon & Christian\\
+Cameron & Daniel & Christian & Evan & Cameron & Daniel\\
+Colby & Jake & Daniel & Evan & Colby & Darwin\\
+Darwin & Jordan & Darwin & Nathan & Evan & Jake\\
+Evan & Nathan & Jake & Timmy & Timmy & Jordan\\
+Tommy & Patrick & Jordan & Tommy & Tommy & Nathan\\
+\hphantom{1} & Timmy & Patrick & \hphantom{1} & \hphantom{1} & Patrick\\
+\hline
+\end{tabular}
+\caption{The partitions arrived at by minimizing each objective function}
+\end{figure}
+
+For each method we define three scoring metrics:
+
+\begin{enumerate}
+\item \textbf{MinCut}.
+This is the traditional metric of a cut on a weighted graph and is the sum of the weights of the edges we would ``cut'' if we were to sever the edges between distinct patrols.
+More formally, for some partition $I, J$ $I \neq J$, the MinCut is defined as:
+$$
+MinCut(I, J) := \sum_{e = (i, j), i \in I, j \in J}^{} c(e)
+$$
+We want this to be as low as possible.
+\item \textbf{FriendCut}.
+This is the sum of the positive, or friendly edge weights within each patrol.
+That is, if $I_+ = \{e=(i,j) | i,j \in I, c(e)=1\}$ is the the set of edges in patrol $I$ with positive weight and $J_+ = \{e=(i,j) | i,j \in J, c(e)=1\}$ is the set of edges in patrol $J$ with positive weight, then (assuming all positive weights are 1):
+$$
+FriendCut(I, J) :=  |I_+ \cup J_+|
+$$
+We want this to be as high as possible.
+\item \textbf{EnemyCut}.
+This is similar to FriendCut, except we now measure the number of negative edges going from one scout to another within the same patrol.
+Let $I_- = \{e=(i,j) | i,j \in I, c(e)=-1\}$ be the the set of edges in patrol $I$ with negative weight and $J_- = \{e=(i,j) | i,j \in J, c(e)=-1\}$ be the set of edges in patrol $J$ with negative weight, then
+$$
+EnemyCut(I, J) := |I_- \cup J_-|
+$$
+    We want this to be as low as possible.
+\end{enumerate}
+
+The other methods are modifications of these three and are defined in Appendix A.
+
+\clearpage
+
+\begin{figure}[h]
+    \centering
+    \begin{tabular}{ |l|l|l|l| }
+        \hline
+        \textbf{Method} & \textbf{MinCut} & \textbf{FriendCut} & \textbf{EnemyCut} \\
+        \hline
+        MinCut & 0 & 17 & 0 \\
+        FriendCut & 1 & 18 & 1 \\
+        EnemyCut & 2 & 15 & 0 \\
+        AwkwardCut (+1) & 0 & 17 & 0 \\
+        AwkwardCut (+2) & 1 & 18 & 2 \\
+        HybridCut & 0 & 18 & 1 \\
+        \hline
+    \end{tabular}
+    \caption{The scores according to the three primary metrics.}
+\end{figure}
+
+Of the objective functions we chose to use, it seems like MinCut and HybridCut are the best performing.
+FriendCut performs almost as well as HybridCut, except for one worse in the EnemyCut objective.
+EnemyCut does poorly since finding a partition with 0 intra-patrol negative weight edges is easy in this example, and so EnemyCut settles for the first suboptimal (in terms of the other objective functions) solution it finds.
+AwkwardCut (+1) arrived at the same partition as MinCut (not altogether surprising, but informative, since even with the additional +1 penalty for placing Brandon and Tommy in the same patrol it decided it would be best off to do so anyways).
+But changing the ``awkward'' penalty to +2, the same penalty as mutual dislike in EnemyCut, we get a completely different partition.
+
+Let's take a more critical look at the optimal partition arrived at via MinCut:
+
+\begin{table}[h]
+	\caption{}\label{eqtable}
+	\renewcommand\arraystretch{1.5}
+	\noindent\[
+	\begin{array}{|c|c|}
+	\hline
+	\textbf{Patrol 1}&\textbf{Patrol 2}\\
+	\hline
+	\text{Brandon, Cameron, Colby,}&\text{Christian, Daniel, Jake,}\\
+	\text{Darwin, Evan, Tommy}&\text{Jordan, Nathan, Patrick, Timmy}\\
+	\hline
+	\end{array}
+	\]
+\end{table}
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.60]{graph}
+\end{figure}
+
+\begin{enumerate}
+	\item Tommy and Timmy both like each other and are in separate patrols.
+But if we move Timmy to Patrol 1 ($P1$), we break up the mutual friendship of Timmy and Nathan in $P2$.
+Suppose we moved both Timmy and Nathan to $P1$ and Brandon (who dislikes Nathan) to $P2$.
+But then Brandon loses two friends (Tommy and Cameron), gains one (Jordan), and now neither Nathan nor Timmy are in the same patrol as Daniel, whom they both like.
+Although, Brandon and Tommy find themselves in the awkward situation of having complete opposite sentiments towards each other.
+Perhaps this would be a worthwhile trade after all (and a larger, negative weight needs to be placed upon disfavor-favor relationships).
+
+	\item As mentioned in the previous point, Tommy dislikes Brandon.
+Trading Tommy in $P1$ for Jordan in $P2$ seems promising, But we have overlooked the fact that Colby and Jordan are our NA pair, and cannot be in the same patrol under any circumstances.
+Trading Colby for Jordan is another option, but Colby will be missed by Darwin, Evan, Cameron and Tommy (In other words, the whole of P1 minus Brandon).
+\end{enumerate}
+
+On the whole, though, there are no obvious improvements that can be made to our algorithm's partition.
+We find that this is an assignment that could have reasonably been arrived at by Scoutmaster Bruce (granted the group dynamics have not changed since the scouts were surveyed).
+Furthermore, the solution took only 10 seconds to arrive at on a single core machine.
+
+\section*{Improvements}
+While our algorithm demonstrably works well on smaller groups of 13 people, we do not expect our brute force solution to be tractable on larger groups wherein greater than 2 partitions are required.
+As a possible scenario, consider a company of 150 people to be divided into fixed-size teams of 10 individuals each.
+There are $150!
+/ (10!)^{15}$ possible partitions, a number proportional to $10^{164}$ - meaning we could instead use that computational time to enumerate the number of atoms in the universe… a tredecillion $(10^{78})$ times.
+Knowing this, perhaps it won’t surprise the reader that this particular graph partitioning problem in general graphs is NP-complete - though there do exist approximation algorithms.%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\footnote{B.
+W.
+Kernighan, S.
+Lin, \textit{An Efficient Heuristic Procedure for Partitioning Graphs}.
+Bell System Technical Journal.
+\textbf{49} (1970), 291--307.
+doi: 10.1002/j.1538-7305.1970.tb01770.x}%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+As scout enrollment increases, it becomes necessary to consider scenarios where we have both a large number of scouts in need of assignment and multiple patrols to choose from.
+Our algorithm could be extended to use brute force to find a globally optimal solution when it is deemed computationally feasible and to use an approximating algorithm to find a locally optimal solution otherwise.
+
+\section*{Conclusions}
+We initially struggled to decide on determining an appropriate objective function to partition the scouts to different groups.
+Follow up interviews with Scoutmaster Gene Bruce allowed us to decide on this function.
+We learned to rely on his expertise and intuition rather than trying to blindly decide on which types of relationships should be fostered and avoided to weight our objective function.
+Thus, we learned necessary communication skills and consideration of our community partner's needs.
+
+With a feasibility check of how many solutions possible once the mathematical model was decided using graphs, a brute force strategy was the most accessible method to arrive at a solution.
+Thus we were able to develop a program that can partition scouts using this strategy.
+We learned to utilize tools we have at hand right away, and then we can consider other methods later on to reduce the computational time of our algorithm.
+
+\section*{Acknowledgments}
+We would like to thank Scoutmaster Gene Bruce for providing us the necessary data and being extremely patient and understanding when discussing the background and problem.
+We would also like to thank Professor Sarah Billey for the mentorship and guidance she provided throughout this project, and our TA Austin Tran for his comments and suggestions.
+
+\section*{Verification Statement}
+(Once we receive our verification statement, we will place it here.)
+
+\section*{Appendix A}
+The other objective functions.
+
+\begin{enumerate}
+    \item \textbf{AwkwardCut (+i)}.
+Same in all respects to MinCut, except for an additional penalty of $i$ if there are two scouts, Scout A and Scout B, in the same patrol such that Scout A likes Scout B and Scout B dislikes Scout A.
+    \begin{equation*}
+        \begin{split}
+            AwkwardCut&(I, J, i) := \Bigg[ \sum_{e = (k, j), k \in I, j \in J}^{} c(e) \Bigg] \\
+            &+ i * \left\vert{\{e_+ = (k,j), e_- = (j, k) : e_+, e_- \in I, c(e_+) = 1, c(e_-) = -1\}}\right\vert
+        \end{split}
+    \end{equation*}
+
+\item \textbf{HybridCut}. A convex combination of MinCut and FriendCut, each with equal weight (that is, half the MinCut loss plus half the FriendCut loss).
+    \begin{equation*}
+        HybridCut(I, J) := \frac{1}{2}MinCut(I, J) + \frac{1}{2}FriendCut(I, J)
+    \end{equation*}
+\end{enumerate}
+
+\bibliographystyle{amsplain}
+\begin{thebibliography}{10}
+
+\bibitem {A} B.
+W.
+Kernighan, S.
+Lin, \textit{An Efficient Heuristic Procedure for Partitioning Graphs}.
+Bell System Technical Journal.
+\textbf{49} (1970), 291--307.
+doi: 10.1002/j.1538-7305.1970.tb01770.x
+
+\end{thebibliography}
+
+\end{document}
+
+%------------------------------------------------------------------------------
+% End of journal.tex
+%------------------------------------------------------------------------------
+=======
+
+\documentclass{amsart}
+
+%     If your article includes graphics, uncomment this command.
+\usepackage{graphicx}
+\usepackage{amsmath,amsfonts,amssymb,amscd,amsthm,amsbsy,upref}
+\usepackage[all]{xy}
+\usepackage{amsmath}
+\usepackage{mathrsfs}
+\usepackage{paralist}
+\usepackage{setspace}
+\usepackage{graphicx}
+\usepackage{tikz-cd}
+\usepackage{pgfplots}
+\usepackage{fancyhdr}
+\usepackage{tikz}
+\usepackage{pgfplots}
+\usepackage{array}
+
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}[theorem]{Lemma}
+
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{example}[theorem]{Example}
+\newtheorem{xca}[theorem]{Exercise}
+
+\theoremstyle{remark}
+\newtheorem{remark}[theorem]{Remark}
+
+\numberwithin{equation}{section}
+
+%    Absolute value notation
+\newcommand{\abs}[1]{\lvert#1\rvert}
+
+%    Blank box placeholder for figures (to avoid requiring any
+%    particular graphics capabilities for printing this document).
+\newcommand{\blankbox}[2]{%
+  \parbox{\columnwidth}{\centering
+%    Set fboxsep to 0 so that the actual size of the box will match the
+%    given measurements more closely.
+    \setlength{\fboxsep}{0pt}%
+    \fbox{\raisebox{0pt}[#2]{\hspace{#1}}}%
+  }%
+}
+
+\pgfplotsset{compat=1.13}
+
+\begin{document}
+
+\title{Assigning Scouts to Optimal Patrols}
+\author{Phil Snyder}
+\author{Ian Bruce}
+\author{Julio Marco Pineda}
+
+\begin{abstract}
+The Scoutmaster wants to find an arrangement of 13 new scouts into two patrols.
+The cohesiveness and overall quality experience of a patrol is severely affected by adverse relationship and can be improved by existing preferences.
+Thus, by using anonymous data from scouts and parents, a brute force strategy was employed to find the optimal arrangement of scouts that avoids severe conflict and maximizes existing preferences.
+An arrangement of 6 and 7 scouts on one patrol respectively was found that satisfies the desired conditions.
+\end{abstract}
+\maketitle
+\section*{Problem Description}
+Every year the Scoutmaster must assign new scouts to groups called patrols.
+In these patrols, scouts would plan and perform different scouting activities together in which they can form close friends that grow together.
+However, in recent years, the number of scouts have been increasing at a rate such that it is no longer logistically viable to place every new scout in a single patrol; multiple patrols are needed to organize activities better and to form proper bonds and community between the new recruits.
+Furthermore, the new scouts are in the fifth or sixth grade who have not fully developed their interpersonal skills.
+Many can be temperamental and awkward when dealing with adversity, and even some devastating conflicts occur where not only the two fighting scouts create animosity between each other, the rest of the troop alienates the destructive pair.
+If enough of these adverse interactions occur within a troop, many scouts would decide to quit and lose out in a great experience and community in the Boy Scouts.
+
+The data the scoutmaster collects about adverse pairings and preferences are either privately given by the scouts themselves or by their parents.
+The possibly destructive interactions are given by the parents of the scouts.Previously, the Scoutmaster was able to determine the perfect patrols manually due to having smaller number of scouts.
+However, the number of scouts have been increasing yearly to the point that solving this grouping problem cannot be done by hand in a reasonable time.
+Therefore, we can help our community partner find a better means to solve his troop organization problem.
+Our goal for this project is to assign boy scouts in different patrols of appropriate size (of 6-8 scouts) such that we maximize the retention rate of the scouts in the program by avoiding severe conflicts while maximizing positive relationships between scouts.
+
+We would like to answer the direct problem our community partner provided of arranging 15 scouts into two troops that avoids severe conflicts and promotes the positive relationships between the scouts.
+Furthermore, we want to determine if we can build a model to predict these patrol arrangements for any number of scouts so that the Scoutmaster can use this model for future years.
+Other than improving the overall experience of the scouts and alleviating the burden of the Scoutmaster, this model can possibly be applied to other situations where arranging a number of people of groups where interpersonal dynamics is necessary and important.
+
+\section*{Simplifications}
+Ultimately, our goal is to find groups for the scouts that will allow them to grow and be happy in the troop.
+If we really wanted to fully understand how a group can foster growth in an individual, we would want to consider the astronomical number of aspects about the personalities of each member and how those influence each of the other members given those other members’ astronomically complex personalities.
+Sociological and psychological literature would need to be studied extensively to learn about these aspects, and experts in the field would most likely be required to collect relevant data for each child, taking several weeks to conduct interviews.
+At the of the day, given our limited knowledge in this area, we settled with considering a subset of the sociological data between the scouts.
+We asked each scout which other scouts they liked and which ones they disliked.
+Now, this is obviously a giant simplification, but the difficulty of gathering more complex information and compiling said information would be too great for the expertise level of anyone involved in solving this problem.
+Also, this process of gathering data would have to be repeated each year for each new batch of new scouts, and would therefore be a burden on the Scoutmaster to collect a lot of data.
+
+Another simplification is to assume that the quality of a group can be determined by the quality of the relationships between the unique pairs of the group - in other words, the whole is equal to the sum of the parts.
+For example, consider the case where scouts A, B and C individually like each other when they are alone with one other scout, but don’t like being in a group together.
+This epiphenomenon won’t be considered; in context of our model, the pairing of these scouts would be heavily favored.
+
+\section*{Mathematical Model}
+The problem of placing individuals into groups visually lends itself well to the notion of a mathematical graph.
+In mathematical vocabulary, a graph consists of a group of objects called vertices that get connected with an "edge" if they share some sort of relationship.
+The graph we will use to represent the data of our problem can be visualized with the image offered on page (5?), where vertices are represented by the circled names and edges are represented by lines between the circles.
+Since there is an inherent sense of relationships between individuals (how much a scout likes another scout), it is appropriate to consider the scouts as vertices in a graph with directed edges representing how much the “head” scout likes the “tail” scout.
+Our goal, using this language, is to find a partition of the vertices in the graph that satisfies some sort of objective function given size constraints of the partitions.
+In our problem at hand, we consider a complete graph with 13 vertices and edge weights being either 0, 1, or -1, corresponding to indifference, approval, or disapproval respectively of the “head” scout with respect to the “tail” scout.
+The Scoutmaster specifically wants two groups out of these 13, and with the size constraints for a patrol considered, our goal becomes finding 6 vertices to be in one partition and the other 7 being in another.
+The total number of partitions becomes 13 choose 6, which is equivalent to 1,716.
+This is certainly within the scope of considering total enumeration of the partitions as a reasonable solution to the problem, therefore we will proceed accordingly and implement the calculation using Python and its associated libraries to represent the data in a format that is convenient to work with.
+
+Now, a natural question for us to consider is what objective function would be most appropriate to optimize.
+A natural answer to this would be to sum the edge weights of all edges that connect vertices that are within partitions.
+This would make sense if separating enemies is just as important as joining friends, but this just might not be the case.
+In terms of weighing these factors, we defer to Scoutmaster Bruce’s experience with how past patrols fared with friends and foes in the mix.
+Therefore, we consider a family of possible objective functions, namely giving real numbered weights to instances of approving relationships and disapproving relationships, and select a few of these objective functions to solve for.
+We then present the collection of results from different objective functions to the Scoutmaster for him to see which one seems to most reasonably match his preferences.
+\section*{Solution of Mathematical Problem}
+We represent our directed graph as a 13 by 13 similarity matrix, or, as might be more fitting for the problem, an “affinity” matrix. Like below:%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\footnote{The scouts were enumerated like so: 1. Brandon, 2. Cameron, 3, Christian, 4. Colby, 5. Daniel, 6. Darwin, 7. Evan, 8. Jake, 9. Jordan, 10. Nathan, 11. Patrick, 12. Timmy, 13. Tommy.}%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\includegraphics[scale=0.6]{solution1.JPG}
+
+The NA value represents a hard constraint established by a scout’s parent (i.e., under no circumstances are these two to be in the same patrol).
+In our implementation we convert the NA value to $-\infty$.
+We designed an objective function that accepts two scouts who are assumed to be in different patrols and calculates the loss associated with such a placement.
+The loss is defined to be $L(x, y) = \text{max}(-2, x[y] + y[x])$ where $x[y]$ is $x$’s disposition towards $y$ and vise versa.
+For example, if $x$ dislikes $y$ and $y$ dislikes $x$, the loss is -2.
+If $x$ likes $y$ but $y$ dislikes $x$ the loss is 0.
+If $x$ and $y$ both like each other, the loss is 2.
+In mixed cases where the first scout is indifferent to another, but the other likes or dislikes the first scout, the loss is +1 or -1 respectively.
+We are trying to minimize the loss, so in applying our loss function over the entire dataset, our algorithm is equally averse to same-patrol mutual enmity as it is to dividing mutual friendship, half as averse to the indifference of one scout and the stronger inclination of another, and indifferent to polar opposite affinities within the same patrol.
+There is a question as to whether this is an accurate reflection of how individual scouts would view their own “loss” of any of the possible scenarios.
+As it turns out, our algorithm arrives at a very reasonable solution regardless.
+\section*{Results}
+As mentioned earlier, the number of scouts we must partition is small enough to find an optimal solution by brute force.
+Our algorithm finds a partition of scouts into patrols of sizes 6 and 7 with a total loss of 0 as described in the table below:
+\begin{table}[ht]
+	\caption{}\label{eqtable}
+	\renewcommand\arraystretch{1.5}
+	\noindent\[
+	\begin{array}{|c|c|}
+	\hline
+	\textbf{Patrol 1}&\textbf{Patrol 2}\\
+	\hline
+	\text{Brandon, Cameron, Colby,}&\text{Christian, Daniel, Jake,}\\
+	\text{Darwin, Evan, Tommy}&\text{Jordan, Nathan, Patrick, Timmy}\\
+	\hline
+	\end{array}
+	\]
+\end{table}
+\includegraphics[scale=0.65]{results1.JPG}
+Of the undesirable outcomes of such a partition, we can spot a few:
+\begin{enumerate}
+	\item Tommy and Timmy both like each other and are in separate patrols.
+But if we move Timmy to Patrol 1 ($P1$), we break up the mutual friendship of Timmy and Nathan in $P2$.
+Suppose we moved both Timmy and Nathan to $P1$ and Brandon (who dislikes Nathan) to $P2$.
+But then Brandon loses two friends (Tommy and Cameron), gains one (Jordan), and now neither Nathan nor Timmy are in the same patrol as Daniel, whom they both like.
+Although, Brandon and Tommy find themselves in the awkward situation of having complete opposite sentiments towards each other.
+Perhaps this would be a worthwhile trade after all (and a larger, negative weight needs to be placed upon disfavor-favor relationships).
+
+	\item As mentioned in the previous point, Tommy dislikes Brandon.
+Trading Tommy in $P1$ for Jordan in $P2$ seems promising, But we have overlooked the fact that Colby and Jordan are our NA pair, and cannot be in the same patrol under any circumstances.
+Trading Colby for Jordan is another option, but Colby will be missed by Darwin, Evan, Cameron and Tommy (In other words, the whole of P1 minus Brandon).
+\end{enumerate}
+On the whole, though, there are no obvious improvements that can be made to our algorithm’s partition.
+We find that this is an assignment that could have reasonably been arrived at by Scoutmaster Bruce (granted the group dynamics have not changed since the scouts were surveyed).
+Furthermore, the solution took only 20 seconds to arrive at on a single core machine.
+
+\section*{Improvements}
+While our algorithm demonstrably works well on smaller groups of 13 people, we do not expect our brute force solution to be tractable on larger groups wherein greater than 2 partitions are required.
+As a possible scenario, consider a company of 150 people to be divided into fixed-size teams of 10 individuals each.
+There are $150!
+/ (10!)^{15}$ possible partitions, a number proportional to $10^164$ - meaning we could instead use that computational time to enumerate the number of atoms in the universe… a tredecillion $(10^{78})$ times.
+Knowing this, perhaps it won’t surprise the reader that this particular graph partitioning problem in general graphs is NP-complete - though there do exist approximation algorithms.%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\footnote{B.
+W.
+Kernighan, S.
+Lin, \textit{An Efficient Heuristic Procedure for Partitioning Graphs}.
+Bell System Technical Journal.
+\textbf{49} (1970), 291--307.
+doi: 10.1002/j.1538-7305.1970.tb01770.x}%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+As scout enrollment increases, it becomes necessary to consider scenarios where we have both a large number of scouts in need of assignment and multiple patrols to choose from.
+Our algorithm could be extended to use brute force to find a globally optimal solution when it is deemed computationally feasible and to use an approximating algorithm to find a locally optimal solution otherwise.
+
+\section*{Conclusions}
+We initially struggled to decide on determining an appropriate objective function to partition the scouts to different groups.
+Follow up interviews with Scoutmaster Gene Bruce allowed us to decide on this function.
+We learned to rely on his expertise and intuition rather than trying to blindly decide on which types of relationships should be fostered and avoided to weight our objective function.
+Thus, we learned necessary communication skills and consideration of our community partner's needs.
+
+With a feasibility check of how many solutions possible once the mathematical model was decided using graphs, a brute force strategy was the most accessible method to arrive at a solution.
+Thus we were able to develop a program that can partition scouts using this strategy.
+We learned to utilize tools we have at hand right away, and then we can consider other methods later on to reduce the computational time of our algorithm.
+
+\section*{Acknowledgments}
+We would like to thank Scoutmaster Gene Bruce for providing us the necessary data and being extremely patient and understanding when discussing the background and problem.
+We would also like to thank Professor Sarah Billey for the mentorship and guidance she provided throughout this project, and our TA Austin Tran for his comments and suggestions.
+
+\section*{Verification Statement}
+(Once we receive our verification statement, we will place it here.)
+
+\bibliographystyle{amsplain}
+\begin{thebibliography}{10}
+
+\bibitem {A} B.
+W.
+Kernighan, S.
+Lin, \textit{An Efficient Heuristic Procedure for Partitioning Graphs}.
+Bell System Technical Journal.
+\textbf{49} (1970), 291--307.
+doi: 10.1002/j.1538-7305.1970.tb01770.x
+
+\end{thebibliography}
+
+\end{document}
+
+%------------------------------------------------------------------------------
+% End of journal.tex
+%------------------------------------------------------------------------------
+>>>>>>> 9771af696fbc57423120850744607162fcf171d5
+=======
 \documentclass{amsart}
@@
 %------------------------------------------------------------------------------
->>>>>>> 9771af696fbc57423120850744607162fcf171d5

### 2016-05-31 poster/poster.tex
poster material added

% Copyright 2004 by Till Tantau <tantau@users.sourceforge.net>.
%
% In principle, this file can be redistributed and/or modified under
% the terms of the GNU Public License, version 2.
%
% However, this file is supposed to be a template to be modified
% for your own needs. For this reason, if you use this file as a
% template and not specifically distribute it as part of a another
% package/program, I grant the extra permission to freely copy and
% modify this file as you see fit and even to delete this copyright
% notice.

\documentclass{beamer}

% There are many different themes available for Beamer. A comprehensive
% list with examples is given here:
% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
% You can uncomment the themes below if you would like to use a different
% one:
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
\usetheme{Boadilla}
%\usetheme{boxes}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{default}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
%\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

\title{Assigning Boy Scouts to Patrols}

% A subtitle is optional and this may be deleted
%\subtitle{Optional Subtitle}

\author{Ian Bruce \and Julio Pineda \and Phil Snyder}

% We can cut the paper here

\date{MATH 480, Community Project}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

\subject{Theoretical Computer Science}
% This is only inserted into the PDF information catalog. Can be left
% out.

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

% \pgfdeclareimage[height=0.5cm]{university-logo}{university-logo-filename}
% \logo{\pgfuseimage{university-logo}}

% Let's get started
\begin{document}

\begin{frame}
  \titlepage
\end{frame}

% Section and subsections will appear in the presentation overview
% and table of contents.
\begin{frame}{The Problem}%{Optional Subtitle}
We need to do a project so we can pass this class.  
\end{frame}

% You can reveal the parts of a slide one at a time
% with the \pause command:
\begin{frame}{A Mathematical Interpretation}
1 enemy + 1 enemy = 0 survivors  
\end{frame}

\begin{frame}{Our Solution}
Isolate unloved scouts to reinforce feelings of social helplessness.
\end{frame}

\begin{frame}{Blocks}
\begin{block}{Block Title}
You can also highlight sections of your presentation in a block, with it's own title
\end{block}
\begin{theorem}
There are separate environments for theorems, examples, definitions and proofs.
\end{theorem}
\begin{example}
Here is an example of an example block.
\end{example}
\end{frame}

% Placing a * after \section means it will not show in the
% outline or table of contents.
\section*{Summary}

\begin{frame}{Summary}
  \begin{itemize}
  \item
    The \alert{first main message} of your talk in one or two lines.
  \item
    The \alert{second main message} of your talk in one or two lines.
  \item
    Perhaps a \alert{third message}, but not more than that.
  \end{itemize}

  \begin{itemize}
  \item
    Outlook
    \begin{itemize}
    \item
      Something you haven't solved.
    \item
      Something else you haven't solved.
    \end{itemize}
  \end{itemize}
\end{frame}

% All of the following is optional and typically not needed. 
\appendix
\section<presentation>*{\appendixname}
\subsection<presentation>*{For Further Reading}

\begin{frame}[allowframebreaks]
  \frametitle<presentation>{For Further Reading}

  \begin{thebibliography}{10}

  \beamertemplatebookbibitems
  % Start with overview books.

  \bibitem{Author1990}
    A.~Author.
    \newblock {\em Handbook of Everything}.
    \newblock Some Press, 1990.

  \beamertemplatearticlebibitems
  % Followed by interesting articles. Keep the list short.

  \bibitem{Someone2000}
    S.~Someone.
    \newblock On this and that.
    \newblock {\em Journal of This and That}, 2(1):50--100,
    2000.
  \end{thebibliography}
\end{frame}

\end{document}

### 2016-06-01 poster/poster.tex
added solution and objective function descriptions

@@
 
+\begin{frame}{The Objective Functions}
+
+\begin{enumerate}
+\scriptsize
+\item \textbf{MinCut}.
+This is the traditional metric of a cut on a weighted graph and is the sum of the weights of the edges we would ``cut'' if we were to sever the edges between distinct patrols.
+More formally, for some partition $I, J$ $I \neq J$, the MinCut is defined as:
+$$
+MinCut(I, J) := \sum_{e = (i, j), i \in I, j \in J}^{} c(e)
+$$
+We want this to be as low as possible.
+\item \textbf{FriendCut}.
+This is the sum of the positive, or friendly edge weights within each patrol.
+That is, if $I_+ = \{e=(i,j) | i,j \in I, c(e)=1\}$ is the the set of edges in patrol $I$ with positive weight and $J_+ = \{e=(i,j) | i,j \in J, c(e)=1\}$ is the set of edges in patrol $J$ with positive weight, then (assuming all positive weights are 1):
+$$
+FriendCut(I, J) :=  |I_+ \cup J_+|
+$$
+We want this to be as high as possible.
+\item \textbf{EnemyCut}.
+This is similar to FriendCut, except we now measure the number of negative edges going from one scout to another within the same patrol.
+Let $I_- = \{e=(i,j) | i,j \in I, c(e)=-1\}$ be the the set of edges in patrol $I$ with negative weight and $J_- = \{e=(i,j) | i,j \in J, c(e)=-1\}$ be the set of edges in patrol $J$ with negative weight, then
+$$
+EnemyCut(I, J) := |I_- \cup J_-|
+$$
+    We want this to be as low as possible.
+\end{enumerate}
+\end{frame}
+
+\begin{frame}
+\begin{enumerate}
+    \scriptsize
+        \setcounter{enumi}{3}
+\item \textbf{AwkwardCut (+i)}.
+Same in all respects to MinCut, except for an additional penalty of $i$ if there are two scouts, Scout A and Scout B, in the same patrol such that Scout A likes Scout B and Scout B dislikes Scout A.
+\begin{equation*}
+    \begin{split}
+        AwkwardCut&(I, J, i) := \Bigg[ \sum_{e = (k, j), k \in I, j \in J}^{} c(e) \Bigg] \\
+        &+ i * \left\vert{\{e_+ = (k,j), e_- = (j, k) : e_+, e_- \in I, c(e_+) = 1, c(e_-) = -1\}}\right\vert
+    \end{split}
+\end{equation*}
+
+\item \textbf{HybridCut}. A convex combination of MinCut and FriendCut, each with equal weight (that is, half the MinCut loss plus half the FriendCut loss).
+\begin{equation*}
+    HybridCut(I, J) := \frac{1}{2}MinCut(I, J) + \frac{1}{2}FriendCut(I, J)
+\end{equation*}
+\end{enumerate}
+\end{frame}
+
 \begin{frame}{Our Solution}
-Isolate unloved scouts to reinforce feelings of social helplessness.
+
+\begin{table}[h]
+\centering
+\footnotesize
+\begin{tabular}{ |l|l|l|l|l|l| }
+\hline
+\multicolumn{2}{|c|}{\textbf{MinCut}} & \multicolumn{2}{|c|}{\textbf{FriendCut}} & \multicolumn{2}{|c|}{\textbf{EnemyCut}}\\
+\hline
+Patrol 1 & Patrol 2 & Patrol 1 & Patrol 2 & Patrol 1 & Patrol 2\\
+\hline
+Brandon & Christian & Brandon & Daniel & Brandon & Christian\\
+Cameron & Daniel & Cameron & Darwin & Cameron & Jake\\
+Colby & Jake & Christian & Jake & Colby & Jordan\\
+Darwin & Jordan & Colby & Jordan  & Daniel & Nathan\\
+Evan & Nathan & Evan & Nathan & Darwin & Patrick\\
+Tommy & Patrick & Tommy & Patrick & Evan & Timmy\\
+& Timmy & & Timmy & Tommy & \\
+\hline
+\multicolumn{6}{|c|}{\hphantom{1}} \\
+\hline
+\multicolumn{2}{|c|}{\textbf{AwkwardCut (+1)}} & \multicolumn{2}{|c|}{\textbf{AwkwardCut (+2)}} & \multicolumn{2}{|c|}{\textbf{HybridCut}} \\
+\hline
+Patrol 1 & Patrol 2 & Patrol 1 & Patrol 2 & Patrol 1 & Patrol 2\\
+\hline
+Brandon & Christian & Brandon & Cameron & Brandon & Christian\\
+Cameron & Daniel & Christian & Evan & Cameron & Daniel\\
+Colby & Jake & Daniel & Evan & Colby & Darwin\\
+Darwin & Jordan & Darwin & Nathan & Evan & Jake\\
+Evan & Nathan & Jake & Timmy & Timmy & Jordan\\
+Tommy & Patrick & Jordan & Tommy & Tommy & Nathan\\
+\hphantom{1} & Timmy & Patrick & \hphantom{1} & \hphantom{1} & Patrick\\
+\hline
+\end{tabular}
+\caption{The partitions arrived at by minimizing each objective function}
+\end{table}
+
 \end{frame}

### 2016-06-01 poster/poster.tex
Problem description slide
Need to finish Simplifications

--- 2016-05-31 poster/poster.tex
@@
 % is a graphic format that can be processed by latex or pdflatex,
-% resp., then you can add a logo as follows:
+% resp., then you can add a logo as follows:x
 
@@
 % and table of contents.
-\begin{frame}{The Problem}%{Optional Subtitle}
-We need to do a project so we can pass this class.  
+\begin{frame}{Problem Description}%{Optional Subtitle}
+	\begin{itemize}
+		\item Scoutmaster Gene Bruce of Troop 407 in Kent, Washington must assign new boy scouts into their patrols. In these patrols, the scouts engage in numerous activities and form close bonds. \\
+		\vspace{0.2in}
+		\item The number of incoming scouts are increasing requiring multiple patrols to better organize activities and to form a small community between scouts.
+		The younger scouts are temperamental and awkward when dealing with adversity, ruining their experience and causing many to quit. \\
+		\vspace{0.2in}
+		\item \textbf{Goal:  Assign 13 boy scouts of appropriate size (of 6-8 scouts) to two patrols such that we maximize the retention rate of the scouts by avoiding severe conflicts and maximizing positive relationships between scouts.}
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Simplifications}
+	\begin{itemize}
+		\item Instead of considering the scouts' complex personality and their interactions with other scouts, we simplified by only asking each scout which other scouts they liked or disliked.
+
+		\item Add last simplification
+	\end{itemize}
 \end{frame}

### 2016-06-01 poster/poster.tex
metrics added
mild changes to prez

--- 2016-06-01 poster/poster.tex
@@
 % is a graphic format that can be processed by latex or pdflatex,
-% resp., then you can add a logo as follows:x
+% resp., then you can add a logo as follows:
 
@@
 % and table of contents.
-\begin{frame}{Problem Description}%{Optional Subtitle}
-	\begin{itemize}
-		\item Scoutmaster Gene Bruce of Troop 407 in Kent, Washington must assign new boy scouts into their patrols. In these patrols, the scouts engage in numerous activities and form close bonds. \\
-		\vspace{0.2in}
-		\item The number of incoming scouts are increasing requiring multiple patrols to better organize activities and to form a small community between scouts.
-		The younger scouts are temperamental and awkward when dealing with adversity, ruining their experience and causing many to quit. \\
-		\vspace{0.2in}
-		\item \textbf{Goal:  Assign 13 boy scouts of appropriate size (of 6-8 scouts) to two patrols such that we maximize the retention rate of the scouts by avoiding severe conflicts and maximizing positive relationships between scouts.}
-	\end{itemize}
-\end{frame}
-
-\begin{frame}{Simplifications}
-	\begin{itemize}
-		\item Instead of considering the scouts' complex personality and their interactions with other scouts, we simplified by only asking each scout which other scouts they liked or disliked.
-
-		\item Add last simplification
-	\end{itemize}
+\begin{frame}{The Problem}%{Optional Subtitle}
+We need to do a project so we can pass this class.  
 \end{frame}
@@
 % with the \pause command:
-\begin{frame}{A Mathematical Interpretation}
-1 enemy + 1 enemy = 0 survivors  
+\begin{frame}{A Mathematical Interpretation - Let's Use Graphs!}
+\begin{definition}
+A \textbf{graph} is a group of objects called vertices where some of the vertices are linked together. When two vertices are linked, we say that there is an \textbf{edge} between them.
+\end{definition}
+Let every scout be a vertex in a graph. Each edge 
 \end{frame}

### 2016-06-01 approximate.py

@@ -1,6 +1,7 @@
 import numpy as np
 import itertools
 import pandas as pd
+import networkx as nx
 
 def rand_affinity(n, probs):
 	return np.random.choice([-1,0,1], (n,n), p=probs)
@@ -19,8 +20,6 @@ def compute_sol(matrix):
 		G = 1
 		Ds = [[diff(i, part) for i in range(len(part))] for j in range(1,3)] # Computes D's
 		
-
-
 	# i - the index of the scout in question
 	# part - the partition scout i is a member of
 	def external(i, part):
@@ -53,5 +52,4 @@ def compute_sol(matrix):
 
 	return optimize_pair([i for i in itertools.chain(itertools.repeat(1, 100), itertools.repeat(2, 100))])
 
-print(compute_sol(mat))
 

### 2016-06-02 poster/poster.tex
Added to poster file
Simplifications and Acknowledgements

--- 2016-06-01 poster/poster.tex
@@
 		\item Instead of considering the scouts' complex personality and their interactions with other scouts, we simplified by only asking each scout which other scouts they liked or disliked.
-
-		\item Add last simplification
+		\vspace{0.5in}
+		\item Quality of the a patrol can be determined by the quality of the relationships between the unique pairs of the group. \\
+		\textbf{i.e.} the whole is equal to the sum of its parts.
 	\end{itemize}
@@
 
+\begin{frame}{Acknowledgments}
+	We would like to sincerely thank:
+	\vspace{0.2in}
+	\begin{itemize}
+		\item \textbf{Scoutmaster Gene Bruce} for providing us the data and being extremely patient and understanding when discussing the background and problem.
+		\vspace{0.2in}
+		\item \textbf{Professor Sara Billey} for the mentorship and guidance throughout the entire project.
+		\vspace{0.2in}
+		\item \textbf{Austin Tran} for his helpful comments and suggestions.
+	\end{itemize}
+\end{frame}
 % You can reveal the parts of a slide one at a time
@@
 
+\begin{frame}{}
+    \begin{block}{How Good Are These Solutions?}
+\begin{figure}[h]
+    \centering
+    \begin{tabular}{ |l|l|l|l| }
+        \hline
+        \textbf{Method} & \textbf{MinCut} & \textbf{FriendCut} & \textbf{EnemyCut} \\
+        \hline
+        MinCut & 0 & 17 & 0 \\
+        FriendCut & 1 & 18 & 1 \\
+        EnemyCut & 2 & 15 & 0 \\
+        AwkwardCut (+1) & 0 & 17 & 0 \\
+        AwkwardCut (+2) & 1 & 18 & 2 \\
+        HybridCut & 0 & 18 & 1 \\
+        \hline
+    \end{tabular}
+    \caption{The scores according to the three primary metrics.}
+\end{figure}
+    \end{block}
+\end{frame}
+
 \begin{frame}{Blocks}

### 2016-06-02 poster/poster.tex

--- 2016-06-01 poster/poster.tex
@@
 \documentclass{beamer}
-
+\usepackage{graphicx}
 % There are many different themes available for Beamer. A comprehensive
@@
 % with the \pause command:
-\begin{frame}{A Mathematical Interpretation}
-1 enemy + 1 enemy = 0 survivors  
+\begin{frame}{A Mathematical Interpretation - Let's Use Graphs!}
+\begin{definition}
+A \textbf{graph} consists of a group of objects called vertices where some of the vertices are linked together. When two vertices are linked, we say that there is an \textbf{edge} between them. 
+\end{definition}
+Graphs are often visualized like this:
+
+Let every scout be a vertex in a graph. Each edge 
 \end{frame}

### 2016-06-02 poster/poster.tex

--- 2016-06-02 poster/poster.tex
@@
 \documentclass{beamer}
-
 % There are many different themes available for Beamer. A comprehensive
@@
 % with the \pause command:
-\begin{frame}{A Mathematical Interpretation}
-1 enemy + 1 enemy = 0 survivors  
+\begin{frame}{A Mathematical Interpretation - Let's Use Graphs!}
+\begin{definition}
+A \textbf{graph} consists of a group of objects called vertices where some of the vertices are linked together. When two vertices are linked, we say that there is an \textbf{edge} between them. 
+\end{definition}
+Implied by the name, graphs have a very visual interpretation, with the following being a typical example of what one would look like:
+\begin{figure}
+\centering
+\includegraphics[scale=.1]{images/graph.png}
+\end{figure}
+The numbered circles represent vertices and the lines between them represent edges.
+We can model a wide variety of real-world things with graphs, namely any collection of objects where clear links can be drawn between pairs that share a relationship
+\end{frame}
+
+\begin{frame}{A Mathematical Interpretation cont.}
+To give a mathematical structure to the patrol problem, we will use what is called a weighted directed graph:
+\begin{definition}
+A graph is \textbf{directed} if edges only represent one way relationships from one vertex to another
+\end{definition}
+\begin{definition}
+A graph is \textbf{weighted} if a numerical value can be assigned to each edge
+\end{definition}
+With these definitions, we can define a graph where each scout is a vertex, and a directed edge from scout A to scout B is defined with weight 1 if scout A likes scout B, 0 if scout A is indifferent to B, and -1 if scout A dislikes scout B.
+\end{frame}
+
+\begin{frame}{A Mathematical Interpretation cont.}
+\begin{itemize}
+\item[]We have our graph, but how can we use it to divide the scouts into patrols? We can split up the vertices into groups that don't overlap, where this splitting is called a \textbf{partition}.
+\item[]Each group would correspond to a potential patrol.
+\item[]We can imagine that there are many different possible partitions for any graph, the question is about which is "best".
+\item[]If we can define some numerical measure of a partition that is best, we can try to \textbf{optimize} that measure of goodness, which manifests as minimizing or maximizing some function dependent on the partition, where we choose to maximize or minimize depending on the measure
+\end{itemize}
 \end{frame}

### 2016-06-03 poster/poster.tex
all the scouts pic

--- 2016-06-02 poster/poster.tex
@@
 		\vspace{0.2in}
-		\item The number of incoming scouts are increasing requiring multiple patrols to better organize activities and to form a small community between scouts.
+		\item The number of incoming scouts are increasing, requiring multiple patrols to better organize activities and to form a small community between scouts.
 		The younger scouts are temperamental and awkward when dealing with adversity, ruining their experience and causing many to quit. \\
@@
 		\item Instead of considering the scouts' complex personality and their interactions with other scouts, we simplified by only asking each scout which other scouts they liked or disliked.
-		\vspace{0.5in}
-		\item Quality of the a patrol can be determined by the quality of the relationships between the unique pairs of the group. \\
-		\textbf{i.e.} the whole is equal to the sum of its parts.
+
+		\item Add last simplification
 	\end{itemize}
@@
 
-\begin{frame}{Acknowledgments}
-	We would like to sincerely thank:
-	\vspace{0.2in}
-	\begin{itemize}
-		\item \textbf{Scoutmaster Gene Bruce} for providing us the data and being extremely patient and understanding when discussing the background and problem.
-		\vspace{0.2in}
-		\item \textbf{Professor Sara Billey} for the mentorship and guidance throughout the entire project.
-		\vspace{0.2in}
-		\item \textbf{Austin Tran} for his helpful comments and suggestions.
-	\end{itemize}
-\end{frame}
 % You can reveal the parts of a slide one at a time

### 2016-06-03 poster/poster.tex
minor

--- 2016-06-02 poster/poster.tex
@@
 \item[]We have our graph, but how can we use it to divide the scouts into patrols? We can split up the vertices into groups that don't overlap, where this splitting is called a \textbf{partition}.
-\item[]Each group would correspond to a potential patrol.
-\item[]We can imagine that there are many different possible partitions for any graph, the question is about which is "best".
-\item[]If we can define some numerical measure of a partition that is best, we can try to \textbf{optimize} that measure of goodness, which manifests as minimizing or maximizing some function dependent on the partition, where we choose to maximize or minimize depending on the measure
+\item[]Each of these groups would correspond to a potential patrol.
+\item[]We can imagine that there would be many different possible ways to split a graph, the key question is figure out which is "best" for the scouts.
+\item[]If we could numerically measure the the "goodness" of a partition, we would want pick the partition that has maximal "goodness".
+Likewise, if we could numerically measure the the "badness" of a partition, we would want pick the partition that has minimal "badness".
+This is the essential method of how to \textbf{optimize} things mathematically.
 \end{itemize}

### 2016-06-06 approximate.py
curr weird state

@@ -6,7 +6,7 @@ import networkx as nx
 def rand_affinity(n, probs):
 	return np.random.choice([-1,0,1], (n,n), p=probs)
 
-mat = rand_affinity(200, [.1, .7, .2])
+mat = rand_affinity(200, [.05, .8, .15]).astype(int)
 np.savetxt("test_mat.csv", mat, delimiter=",")
 
 # In this algorithm, we will represent partitions as incidence vectors, with 1
@@ -19,6 +19,7 @@ def compute_sol(matrix):
 	def optimize_pair(part):
 		G = 1
 		Ds = [[diff(i, part) for i in range(len(part))] for j in range(1,3)] # Computes D's
+		return Ds
 		
 	# i - the index of the scout in question
 	# part - the partition scout i is a member of
@@ -52,4 +53,4 @@ def compute_sol(matrix):
 
 	return optimize_pair([i for i in itertools.chain(itertools.repeat(1, 100), itertools.repeat(2, 100))])
 
-
+print(compute_sol(mat))

### 2016-06-06 draft1.tex
metrics added

--- 2016-05-25 draft1.tex
@@
 
-We aim to answer the direct problem our community partner provided of arranging 13 scouts into two troops that avoids severe conflicts and promotes the positive relationships between the scouts.
+We aim to answer the specific problem our community partner, that of arranging 13 scouts into two troops that avoids severe conflicts and promotes positive relationships between the scouts.
 Furthermore, we want to determine if we can build a model to predict these patrol arrangements for any number of scouts so that the Scoutmaster can use this model for future years.
@@
 \section*{Simplifications}
-Ultimately, our goal is to find groups for the scouts that will allow them to grow and be happy in the troop.
-If we really wanted to fully understand how a group can foster growth in an individual, we would want to consider the astronomical number of aspects about the personalities of each member and how those influence each of the other members given those other members’ astronomically complex personalities.
-Sociological and psychological literature would need to be studied extensively to learn about these aspects, and experts in the field would most likely be required to collect relevant data for each child, taking several weeks to conduct interviews.
-At the of the day, given our limited knowledge in this area, we settled with considering a subset of the sociological data between the scouts.
+Ultimately, our goal is to find groups for the scouts that will allow them to enjoy their time scouting to the fullest.
 We asked each scout which other scouts they liked and which ones they disliked.
-Now, this is obviously a giant simplification, but the difficulty of gathering more complex information and compiling said information would be too great for the expertise level of anyone involved in solving this problem.
-Also, this process of gathering data would have to be repeated each year for each new batch of new scouts, and would therefore be a burden on the Scoutmaster to collect a lot of data.
+Of couse, this is a simplification, but the difficulty in modeling the intracicies of human relationships demands it.
+Besides, our process will have to be repeated each year for every new batch of scouts, and would perhaps be too burdensome to collect to warrant the effort.
 
@@
 As it turns out, our algorithm arrives at a very reasonable solution regardless.
+
 \section*{Results}
@@
 %------------------------------------------------------------------------------
-=======
-
-\documentclass{amsart}
-
-%     If your article includes graphics, uncomment this command.
-\usepackage{graphicx}
-\usepackage{amsmath,amsfonts,amssymb,amscd,amsthm,amsbsy,upref}
-\usepackage[all]{xy}
-\usepackage{amsmath}
-\usepackage{mathrsfs}
-\usepackage{paralist}
-\usepackage{setspace}
-\usepackage{graphicx}
-\usepackage{tikz-cd}
-\usepackage{pgfplots}
-\usepackage{fancyhdr}
-\usepackage{tikz}
-\usepackage{pgfplots}
-\usepackage{array}
-
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}[theorem]{Example}
-\newtheorem{xca}[theorem]{Exercise}
-
-\theoremstyle{remark}
-\newtheorem{remark}[theorem]{Remark}
-
-\numberwithin{equation}{section}
-
-%    Absolute value notation
-\newcommand{\abs}[1]{\lvert#1\rvert}
-
-%    Blank box placeholder for figures (to avoid requiring any
-%    particular graphics capabilities for printing this document).
-\newcommand{\blankbox}[2]{%
-  \parbox{\columnwidth}{\centering
-%    Set fboxsep to 0 so that the actual size of the box will match the
-%    given measurements more closely.
-    \setlength{\fboxsep}{0pt}%
-    \fbox{\raisebox{0pt}[#2]{\hspace{#1}}}%
-  }%
-}
-
-\pgfplotsset{compat=1.13}
-
-\begin{document}
-
-\title{Assigning Scouts to Optimal Patrols}
-\author{Phil Snyder}
-\author{Ian Bruce}
-\author{Julio Marco Pineda}
-
-\begin{abstract}
-The Scoutmaster wants to find an arrangement of 13 new scouts into two patrols.
-The cohesiveness and overall quality experience of a patrol is severely affected by adverse relationship and can be improved by existing preferences.
-Thus, by using anonymous data from scouts and parents, a brute force strategy was employed to find the optimal arrangement of scouts that avoids severe conflict and maximizes existing preferences.
-An arrangement of 6 and 7 scouts on one patrol respectively was found that satisfies the desired conditions.
-\end{abstract}
-\maketitle
-\section*{Problem Description}
-Every year the Scoutmaster must assign new scouts to groups called patrols.
-In these patrols, scouts would plan and perform different scouting activities together in which they can form close friends that grow together.
-However, in recent years, the number of scouts have been increasing at a rate such that it is no longer logistically viable to place every new scout in a single patrol; multiple patrols are needed to organize activities better and to form proper bonds and community between the new recruits.
-Furthermore, the new scouts are in the fifth or sixth grade who have not fully developed their interpersonal skills.
-Many can be temperamental and awkward when dealing with adversity, and even some devastating conflicts occur where not only the two fighting scouts create animosity between each other, the rest of the troop alienates the destructive pair.
-If enough of these adverse interactions occur within a troop, many scouts would decide to quit and lose out in a great experience and community in the Boy Scouts.
-
-The data the scoutmaster collects about adverse pairings and preferences are either privately given by the scouts themselves or by their parents.
-The possibly destructive interactions are given by the parents of the scouts.Previously, the Scoutmaster was able to determine the perfect patrols manually due to having smaller number of scouts.
-However, the number of scouts have been increasing yearly to the point that solving this grouping problem cannot be done by hand in a reasonable time.
-Therefore, we can help our community partner find a better means to solve his troop organization problem.
-Our goal for this project is to assign boy scouts in different patrols of appropriate size (of 6-8 scouts) such that we maximize the retention rate of the scouts in the program by avoiding severe conflicts while maximizing positive relationships between scouts.
-
-We would like to answer the direct problem our community partner provided of arranging 15 scouts into two troops that avoids severe conflicts and promotes the positive relationships between the scouts.
-Furthermore, we want to determine if we can build a model to predict these patrol arrangements for any number of scouts so that the Scoutmaster can use this model for future years.
-Other than improving the overall experience of the scouts and alleviating the burden of the Scoutmaster, this model can possibly be applied to other situations where arranging a number of people of groups where interpersonal dynamics is necessary and important.
-
-\section*{Simplifications}
-Ultimately, our goal is to find groups for the scouts that will allow them to grow and be happy in the troop.
-If we really wanted to fully understand how a group can foster growth in an individual, we would want to consider the astronomical number of aspects about the personalities of each member and how those influence each of the other members given those other members’ astronomically complex personalities.
-Sociological and psychological literature would need to be studied extensively to learn about these aspects, and experts in the field would most likely be required to collect relevant data for each child, taking several weeks to conduct interviews.
-At the of the day, given our limited knowledge in this area, we settled with considering a subset of the sociological data between the scouts.
-We asked each scout which other scouts they liked and which ones they disliked.
-Now, this is obviously a giant simplification, but the difficulty of gathering more complex information and compiling said information would be too great for the expertise level of anyone involved in solving this problem.
-Also, this process of gathering data would have to be repeated each year for each new batch of new scouts, and would therefore be a burden on the Scoutmaster to collect a lot of data.
-
-Another simplification is to assume that the quality of a group can be determined by the quality of the relationships between the unique pairs of the group - in other words, the whole is equal to the sum of the parts.
-For example, consider the case where scouts A, B and C individually like each other when they are alone with one other scout, but don’t like being in a group together.
-This epiphenomenon won’t be considered; in context of our model, the pairing of these scouts would be heavily favored.
-
-\section*{Mathematical Model}
-This problem visually lends itself well to the notion of a mathematical graph.
-Since there is an inherent sense of relationships between individuals (how much a scout likes another scout), it is appropriate to consider the scouts as vertices in a graph with directed edges representing how much the “head” scout likes the “tail” scout.
-Our goal, using this language, is to find a partition of the vertices in the graph that satisfies some sort of objective function given size constraints of the partitions.
-In our problem at hand, we consider a complete graph with 13 vertices and edge weights being either 0, 1, or -1, corresponding to indifference, approval, or disapproval respectively of the “head” scout with respect to the “tail” scout.
-The Scoutmaster specifically wants two groups out of these 13, and with the size constraints for a patrol considered, our goal becomes finding 6 vertices to be in one partition and the other 7 being in another.
-The total number of partitions becomes 13 choose 6, which is equivalent to 1,716.
-This is certainly within the scope of considering total enumeration of the partitions as a reasonable solution to the problem, therefore we will proceed accordingly and implement the calculation using Python and its associated libraries to represent the data in a format that is convenient to work with.
-
-Now, a natural question for us to consider is what objective function would be most appropriate to optimize.
-A natural answer to this would be to sum the edge weights of all edges that connect vertices that are within partitions.
-This would make sense if separating enemies is just as important as joining friends, but this just might not be the case.
-In terms of weighing these factors, we defer to Scoutmaster Bruce’s experience with how past patrols fared with friends and foes in the mix.
-Therefore, we consider a family of possible objective functions, namely giving real numbered weights to instances of approving relationships and disapproving relationships, and select a few of these objective functions to solve for.
-We then present the collection of results from different objective functions to the Scoutmaster for him to see which one seems to most reasonably match his preferences.
-\section*{Solution of Mathematical Problem}
-We represent our directed graph as a 13 by 13 similarity matrix, or, as might be more fitting for the problem, an “affinity” matrix. Like below:%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\footnote{The scouts were enumerated like so: 1. Brandon, 2. Cameron, 3, Christian, 4. Colby, 5. Daniel, 6. Darwin, 7. Evan, 8. Jake, 9. Jordan, 10. Nathan, 11. Patrick, 12. Timmy, 13. Tommy.}%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\includegraphics[scale=0.6]{solution1.JPG}
-
-The NA value represents a hard constraint established by a scout’s parent (i.e., under no circumstances are these two to be in the same patrol).
-In our implementation we convert the NA value to $-\infty$.
-We designed an objective function that accepts two scouts who are assumed to be in different patrols and calculates the loss associated with such a placement.
-The loss is defined to be $L(x, y) = \text{max}(-2, x[y] + y[x])$ where $x[y]$ is $x$’s disposition towards $y$ and vise versa.
-For example, if $x$ dislikes $y$ and $y$ dislikes $x$, the loss is -2.
-If $x$ likes $y$ but $y$ dislikes $x$ the loss is 0.
-If $x$ and $y$ both like each other, the loss is 2.
-In mixed cases where the first scout is indifferent to another, but the other likes or dislikes the first scout, the loss is +1 or -1 respectively.
-We are trying to minimize the loss, so in applying our loss function over the entire dataset, our algorithm is equally averse to same-patrol mutual enmity as it is to dividing mutual friendship, half as averse to the indifference of one scout and the stronger inclination of another, and indifferent to polar opposite affinities within the same patrol.
-There is a question as to whether this is an accurate reflection of how individual scouts would view their own “loss” of any of the possible scenarios.
-As it turns out, our algorithm arrives at a very reasonable solution regardless.
-\section*{Results}
-As mentioned earlier, the number of scouts we must partition is small enough to find an optimal solution by brute force.
-Our algorithm finds a partition of scouts into patrols of sizes 6 and 7 with a total loss of 0 as described in the table below:
-\begin{table}[ht]
-	\caption{}\label{eqtable}
-	\renewcommand\arraystretch{1.5}
-	\noindent\[
-	\begin{array}{|c|c|}
-	\hline
-	\textbf{Patrol 1}&\textbf{Patrol 2}\\
-	\hline
-	\text{Brandon, Cameron, Colby,}&\text{Christian, Daniel, Jake,}\\
-	\text{Darwin, Evan, Tommy}&\text{Jordan, Nathan, Patrick, Timmy}\\
-	\hline
-	\end{array}
-	\]
-\end{table}
-\includegraphics[scale=0.65]{results1.JPG}
-Of the undesirable outcomes of such a partition, we can spot a few:
-\begin{enumerate}
-	\item Tommy and Timmy both like each other and are in separate patrols.
-But if we move Timmy to Patrol 1 ($P1$), we break up the mutual friendship of Timmy and Nathan in $P2$.
-Suppose we moved both Timmy and Nathan to $P1$ and Brandon (who dislikes Nathan) to $P2$.
-But then Brandon loses two friends (Tommy and Cameron), gains one (Jordan), and now neither Nathan nor Timmy are in the same patrol as Daniel, whom they both like.
-Although, Brandon and Tommy find themselves in the awkward situation of having complete opposite sentiments towards each other.
-Perhaps this would be a worthwhile trade after all (and a larger, negative weight needs to be placed upon disfavor-favor relationships).
-
-	\item As mentioned in the previous point, Tommy dislikes Brandon.
-Trading Tommy in $P1$ for Jordan in $P2$ seems promising, But we have overlooked the fact that Colby and Jordan are our NA pair, and cannot be in the same patrol under any circumstances.
-Trading Colby for Jordan is another option, but Colby will be missed by Darwin, Evan, Cameron and Tommy (In other words, the whole of P1 minus Brandon).
-\end{enumerate}
-On the whole, though, there are no obvious improvements that can be made to our algorithm’s partition.
-We find that this is an assignment that could have reasonably been arrived at by Scoutmaster Bruce (granted the group dynamics have not changed since the scouts were surveyed).
-Furthermore, the solution took only 20 seconds to arrive at on a single core machine.
-
-\section*{Improvements}
-While our algorithm demonstrably works well on smaller groups of 13 people, we do not expect our brute force solution to be tractable on larger groups wherein greater than 2 partitions are required.
-As a possible scenario, consider a company of 150 people to be divided into fixed-size teams of 10 individuals each.
-There are $150!
-/ (10!)^{15}$ possible partitions, a number proportional to $10^164$ - meaning we could instead use that computational time to enumerate the number of atoms in the universe… a tredecillion $(10^{78})$ times.
-Knowing this, perhaps it won’t surprise the reader that this particular graph partitioning problem in general graphs is NP-complete - though there do exist approximation algorithms.%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\footnote{B.
-W.
-Kernighan, S.
-Lin, \textit{An Efficient Heuristic Procedure for Partitioning Graphs}.
-Bell System Technical Journal.
-\textbf{49} (1970), 291--307.
-doi: 10.1002/j.1538-7305.1970.tb01770.x}%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-As scout enrollment increases, it becomes necessary to consider scenarios where we have both a large number of scouts in need of assignment and multiple patrols to choose from.
-Our algorithm could be extended to use brute force to find a globally optimal solution when it is deemed computationally feasible and to use an approximating algorithm to find a locally optimal solution otherwise.
-
-\section*{Conclusions}
-We initially struggled deciding on an appropriate objective function to partition the scouts to different groups.
-Follow up interviews with Scoutmaster Gene Bruce allowed us to decide on this function.
-We relied on his expertise and intuition rather than trying to blindly decide on which types of relationships should be fostered and avoided to determine our objective function.
-Thus, we learned necessary communication skills and consideration of our community partner's needs.
-
-With a feasibility check of how many solutions possible once the mathematical model was decided using graphs, a brute force strategy was the most reliable method to arrive at a solution.
-Thus we were able to develop a program that can partition scouts using this strategy.
-We learned to utilize tools we have at hand.
-Further consideration of other methods later on to reduce the computational time of our algorithm.
-
-\section*{Acknowledgments}
-We would like to thank Scoutmaster Gene Bruce for providing us the necessary data and being extremely patient and understanding when discussing the background and problem.
-We would also like to thank Professor Sara Billey for the mentorship and guidance she provided throughout this project, and our TA Austin Tran for his comments and suggestions.
-
-\section*{Verification Statement}
-(Once we receive our verification statement, we will place it here.)
-
-\bibliographystyle{amsplain}
-\begin{thebibliography}{10}
-
-\bibitem {A} B.
-W.
-Kernighan, S.
-Lin, \textit{An Efficient Heuristic Procedure for Partitioning Graphs}.
-Bell System Technical Journal.
-\textbf{49} (1970), 291--307.
-doi: 10.1002/j.1538-7305.1970.tb01770.x
-
-\end{thebibliography}
-
-\end{document}
-
-%------------------------------------------------------------------------------
-% End of journal.tex
-%------------------------------------------------------------------------------
->>>>>>> 9771af696fbc57423120850744607162fcf171d5

### 2016-06-07 draft1.tex
revisions of most sections

@@
 The cohesiveness and overall quality experience of a patrol is affected by adverse relationships and existing friendships.
-Thus, by using anonymous data from scouts and parents, a brute force strategy was employed to find the optimal arrangement of scouts that avoids severe conflict and maximizes existing preferences.
-We found an optimal arrangement of the 13 boy scouts that satisfies the desired conditions.
+By using anonymous data from scouts and parents, a brute force strategy was employed to find an optimal arrangement of scouts that avoids severe conflict while maximizing pre-existing preferences.
 \end{abstract}
+
 \maketitle
@@
 The community partner we are serving for this project is Scoutmaster Gene Bruce of the Boys Scouts of America.
-He leads Troop 407 in Kent, Washington and every year he must assign new scouts to groups called patrols.
-In these patrols, scouts would plan, perform different scouting activities together and form close bonds.
-However, in recent years, the number of scouts have been increasing at such a rate that it is no longer logistically viable to place every new scout in a single patrol; multiple patrols are needed to organize activities better and to form a community between the new recruits.
-Furthermore, the new scouts are in the fifth or sixth grade and have not fully developed their interpersonal skills.
-Many can be temperamental and awkward when dealing with adversity, and even some devastating conflicts occur where not only the two fighting scouts create animosity between each other, the rest of the troop alienates the destructive pair.
-If enough of these adverse interactions occur within a troop, many scouts would decide to quit and lose out in a great experience and community in the Boy Scouts.
+He leads Troop 407 in Kent, Washington and every year he must assign new scouts to groups - called ``patrols''.
+In these patrols, scouts perform different scouting activities together and form close bonds.
+However, in recent years, the number of scouts have been increasing at such a rate that it is no longer logistically viable to place every new scout in a single patrol; multiple patrols are needed to organize activities better and to form a community between new recruits.
+Furthermore, scouts are in the fifth or sixth grade and have not yet fully developed their interpersonal skills.
+Many can be temperamental and awkward when dealing with adversity.
+Devastating conflicts may occur if two at-odds scouts create animosity between each other, alienating themselves from the rest of the patrol.
+If enough of these adverse interactions occur within a patrol, scouts may decide to drop out of the Boy Scouts, losing out on a potentially great experience and community of friends.
 
-The data the Scoutmaster collects about adverse pairings and preferences are either privately given by the scouts themselves.
-Possibly destructive interactions are given by the parents.
+The data the Scoutmaster collects about adverse pairings and preferences are either privately given by the scouts themselves or by their parents.
+The parents preferences are hard rules that we must abide by (if possible).
 In the past, the Scoutmaster determined the perfect arrangement of patrols without too much time and effort due to the smaller number of new recruits.
-However, newer scouts are joining yearly to the point that solving this grouping problem cannot be done by hand in a reasonable time.
+However, newer scouts are joining yearly to the point that solving this grouping problem cannot be done by hand in a reasonable amount of time.
 Therefore, we can help our community partner find a better means to solve his troop organization problem.
-Our goal for this project is to assign boy scouts in different patrols of appropriate size (of 6-8 scouts) such that we maximize the retention rate of the scouts in the program by avoiding severe conflicts while maximizing positive relationships between scouts.
+Our goal is to assign boy scouts in different patrols of appropriate size (of 6-8 scouts) such that we maximize the retention rate of the scouts in the program by avoiding severe conflicts while maximizing positive relationships between scouts.
 
-We aim to answer the specific problem our community partner, that of arranging 13 scouts into two troops that avoids severe conflicts and promotes positive relationships between the scouts.
-Furthermore, we want to determine if we can build a model to predict these patrol arrangements for any number of scouts so that the Scoutmaster can use this model for future years.
-Additionally, this model can possibly be applied to other situations in which arranging a number of people into groups where interpersonal dynamics is necessary and important such as project teams in a company and forming new military troops.
-
-\section*{Data}
+We aim to answer the particular instance of this problem provided by our community partner, that of arranging 13 scouts into two troops that avoids severe conflicts while promoting positive relationships between scouts.
+Furthermore, we want to determine if we can build a model to determine optimal patrol arrangements for any number of scouts so that the Scoutmaster can use this model for future years.
+For 13 scouts, a grouping into one patrol is too large, while three patrols are too small, and need not be considered.
+But in general it is possible to calculate all possible partitions with size constraints on the partitions for an arbitrary number of scouts\footnote{Desquelles, P.
+"Calculation of the Number of Partitions with Constraints on the Fragment Size." Phys.
+Rev.
+C Physical Review C 65.3 (2002)}.
+Additionally, this model could be applied to other situations in which arranging a number of individuals into groups where interpersonal dynamics are necessary to consider - such as teams within a company or military troops.
 The 13 different recruited scouts are Brandon, Cameron, Christian, Colby, Daniel, Darwin, Evan, Jake, Jordan, Nathan, Patrick, Timmy and Tommy.
@@
 For example, there is a dashed arrow from Tommy to Jordan showing that Tommy does not want to be with Jordan.
-Furthermore, the Scoutmaster provided us a separate email, and indicated that Colby and Jordan cannot be in the same troop.
-The highlighted arrows in the figure above does not mean anything new between the relationships of the boy scouts.
+Furthermore, the Scoutmaster provided us a separate email, and indicated that Colby and Jordan cannot be in the same troop (as per a parental request).
+The highlighted arrows in the figure indicate a mutual friendship between scouts.
 
-The Stirling number of the second kind provides the total possible ways to arrange any number of distinguishable objects to indistinguishable groups as long as any troop size is permitted and no possible empty troops.
-For the data in this project, the Stirling number of the second kind for 13 scouts into 2 troops is 4095.
+The Stirling number of the second kind provides the total possible ways to arrange any number of distinguishable objects to indistinguishable groups as long as any troop size is permitted (except the empty troop).
+For the data in this project, the Stirling number of the second kind for 13 scouts into 2 troops is 4095\footnote{See Appendix B}.
 This number provides an upper bound on the possible arrangements for the specific problem tackled in this project.
-
-\section*{Previous Work}
-(Currently reading more on Max Flow-Min cut, K-means clustering, Graph clustering: http://www.sciencedirect.com/science/article/pii/S1574013707000020. We will talk about generating random solutions)
 
@@
 Ultimately, our goal is to find groups for the scouts that will allow them to enjoy their time scouting to the fullest.
-We asked each scout which other scouts they liked and which ones they disliked.
-Of couse, this is a simplification, but the difficulty in modeling the intracicies of human relationships demands it.
-Besides, our process will have to be repeated each year for every new batch of scouts, and would perhaps be too burdensome to collect to warrant the effort.
+Modeling interpersonal relations by likes and dislikes is, of couse, a simplification.
+But the difficulty in modeling the intricacies of human relationships demands it.
+In addition, our process will have to be repeated each year for every new batch of scouts, and would perhaps be too burdensome to warrant the effort if our process attempted to delve too far into the psyche.
 
-Another simplification is to assume that the quality of a group can be determined by the quality of the relationships between the unique pairs of the group - in other words, the whole is equal to the sum of the parts.
-For example, consider the case where scouts A, B and C individually like each other when they are alone with one other scout, but don’t like being in a group together.
-This epiphenomenon won’t be considered; in context of our model, the pairing of these scouts would be heavily favored.
+Another simplification is to assume that the quality of a group can be determined by the quality of relationships between unique pairs either inside or outside the group - in other words, the whole is equal to the sum of the parts.
+For example, consider the case where scouts A, B and C individually like each other when they are alone with one of their group, but don't like being in a group together.
+This type of epiphenomenon won't be considered.
 
@@
 A directed edge between Scout A and Scout B has a weight proportional to how much Scout A likes Scout B.
-We would then like to find a partition of the vertices in the graph that minimizes an objective function given constraints on the sizes of subsets generated by our partition.
-This results in a complete graph on 13 vertices having edge weights 0, 1, or -1, corresponding to indifference, approval, or disapproval, respectively.
+We would then like to find a partition of the vertices in the graph that either minimizes or maximizes an objective function given constraints on the sizes of subsets generated by our partition.
+This results in a complete, directed graph on 13 vertices having edge weights 0, 1, or -1, corresponding to indifference, approval, or disapproval, respectively.
 Scoutmaster Bruce specifically would like two patrols from the 13 scouts - one of size 6, the other of size 7.
 The total number of partitions given these constrants is $\binom{13}{6} = \binom{13}{7} = 1716$.
-This number is reasonable enough that we may calculate the total loss of every valid grouping and select the partition with the lowest score.
+This number is reasonable enough that we may calculate the total loss of every valid grouping and select the partition with the lowest or highest score depending on the objective functionn.
 
@@
 EnemyCut is meant to minimize the number of intra-patrol negative edge weights, Awkward cut ($+i$) is the same as MinCut, but with an additional penalty of $i$ if there are a pair of scouts, Scout A and Scout B, such that Scout A favors Scout B, but Scout B disfavors Scout A.
-Hybrid Cut is a convex combination of MinCut and FriendCut, each with equal weight (that is, half the MinCut loss plus half the FriendCut loss).
+HybridCut is a convex combination of MinCut and FriendCut, each with equal weight (that is, half the MinCut loss plus half the FriendCut loss).
 Our three primary objective functions - MinCut, FriendCut, and EnemyCut are given a more rigorous treatment in the next section.
@@
 We designed a number of objective functions that accept two scouts and calculates the loss associated with such a placement.
-The loss is defined to be $L(x, y) = \text{max}(-2, x[y] + y[x])$ where $x[y]$ is $x$'s disposition towards $y$ and vise versa.
-For example, if $x$ dislikes $y$ and $y$ dislikes $x$, the loss is -2.
-If $x$ likes $y$ but $y$ dislikes $x$ the loss is 0.
-If $x$ and $y$ both like each other, the loss is 2.
-In mixed cases where the first scout is indifferent to another, but the other likes or dislikes the first scout, the loss is +1 or -1 respectively.
-We are trying to minimize the loss, so in applying our loss function over the entire dataset, our algorithm is equally averse to same-patrol mutual enmity as it is to dividing mutual friendship, half as averse to the indifference of one scout and the stronger inclination of another, and indifferent to polar opposite affinities within the same patrol.
-There is a question as to whether this is an accurate reflection of how individual scouts would view their own ``loss'' of any of the possible scenarios.
-As it turns out, our algorithm arrives at a very reasonable solution regardless.
 
@@
 Our algorithm finds a partition of scouts into patrols of sizes 6 and 7 which is optimal with respect to each objective.
-
-\clearpage
 
@@
 Cameron & Daniel & Christian & Evan & Cameron & Daniel\\
-Colby & Jake & Daniel & Evan & Colby & Darwin\\
+Colby & Jake & Daniel & Colby & Colby & Darwin\\
 Darwin & Jordan & Darwin & Nathan & Evan & Jake\\
@@
 \end{figure}
+
+\clearpage
 
@@
 
-\clearpage
-
 \begin{figure}[h]
@@
 
+\clearpage
+
 Of the objective functions we chose to use, it seems like MinCut and HybridCut are the best performing.
@@
 \begin{table}[h]
-	\caption{}\label{eqtable}
+	\caption{The partition arrived at by optimizing MinCut}\label{eqtable}
 	\renewcommand\arraystretch{1.5}
@@
 
-On the whole, though, there are no obvious improvements that can be made to our algorithm's partition.
+On the whole, though, there are no obvious improvements that can be made to our algorithm's partitions.
 We find that this is an assignment that could have reasonably been arrived at by Scoutmaster Bruce (granted the group dynamics have not changed since the scouts were surveyed).
@@
 There are $150!
-/ (10!)^{15}$ possible partitions, a number proportional to $10^{164}$ - meaning we could instead use that computational time to enumerate the number of atoms in the universe… a tredecillion $(10^{78})$ times.
-Knowing this, perhaps it won’t surprise the reader that this particular graph partitioning problem in general graphs is NP-complete - though there do exist approximation algorithms.%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\footnote{B.
+/ (10!)^{15}$ possible partitions, a number proportional to $10^{164}$ - meaning we could instead use that computational time to enumerate the number of atoms in the universe\ldots a tredecillion $(10^{78})$ times.
+Knowing this, perhaps it won't surprise the reader that this particular graph partitioning problem in general graphs is NP-complete - though there do exist approximation algorithms\footnote{B.
 W.
@@
 \textbf{49} (1970), 291--307.
-doi: 10.1002/j.1538-7305.1970.tb01770.x}%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+doi: 10.1002/j.1538-7305.1970.tb01770.x}.
 As scout enrollment increases, it becomes necessary to consider scenarios where we have both a large number of scouts in need of assignment and multiple patrols to choose from.
@@
 \section*{Conclusions}
-We initially struggled to decide on determining an appropriate objective function to partition the scouts to different groups.
-Follow up interviews with Scoutmaster Gene Bruce allowed us to decide on this function.
-We learned to rely on his expertise and intuition rather than trying to blindly decide on which types of relationships should be fostered and avoided to weight our objective function.
-Thus, we learned necessary communication skills and consideration of our community partner's needs.
+Assigning individuals to compatible groups is hard (computationally speaking).
+When the group is small enough - as it is here - globally optimal solutions may be arrived at by checking all feasible solutions.
+But the amount of computation time needed to arrive at an answer grows exponentially with the number of individuals and groups.
+For example, finding an optimal partition of 13 scouts into \textit{three} groups would take over 50 times as long as finding the optimal partition of 13 scouts into two groups (90,090 feasible combinations to check versus 1,716 in the two patrol case).
 
-With a feasibility check of how many solutions possible once the mathematical model was decided using graphs, a brute force strategy was the most accessible method to arrive at a solution.
-Thus we were able to develop a program that can partition scouts using this strategy.
-We learned to utilize tools we have at hand right away, and then we can consider other methods later on to reduce the computational time of our algorithm.
+Nevertheless, our chosen objective functions seem to arrive at solutions that are able to hold up to human scrutiny, and perform well enough when total enumeration is possible.
+Larger problems may require approximating or clustering methods\footnote{Hartigan, J.
+A., and M.
+A.
+Wong.
+"Algorithm AS 136: A K-Means Clustering Algorithm." Applied Statistics 28.1 (1979): 100.}.
 
 \section*{Acknowledgments}
-We would like to thank Scoutmaster Gene Bruce for providing us the necessary data and being extremely patient and understanding when discussing the background and problem.
-We would also like to thank Professor Sarah Billey for the mentorship and guidance she provided throughout this project, and our TA Austin Tran for his comments and suggestions.
+We would like to thank Scoutmaster Gene Bruce for his cordiality and for providing us the necessary data.
+We would also like to thank Professor Sara Billey for her mentorship and guidance provided throughout our project.
 
@@
 
+\bibitem {B} Desquelles, P. "Calculation of the Number of Partitions with Constraints on the Fragment Size." Phys. Rev. C Physical Review C 65.3 (2002)
+
+\bibitem {C} Hartigan, J. A., and M. A. Wong. "Algorithm AS 136: A K-Means Clustering Algorithm." Applied Statistics 28.1 (1979): 100.
+
 \end{thebibliography}

### 2016-06-08 draft1.tex
General edits
Added some details

@@
 Scoutmaster Gene Bruce of the Boy Scouts of America for Troop 407 in Kent, Washington wants to find an arrangement of 13 new scouts into two patrols.
-The cohesiveness and overall quality experience of a patrol is affected by adverse relationships and existing friendships.
-By using anonymous data from scouts and parents, a brute force strategy was employed to find an optimal arrangement of scouts that avoids severe conflict while maximizing pre-existing preferences.
+The cohesiveness and overall quality experience of a patrol is affected by conflicting relationships and existing friendships.
+If enough adverse encounters occur in one patrol, the scouts may decide to drop out from the organization all together, missing out on a potentially great experience and community of friends.
+Thus our project focuses on helping the Scoutmaster find an optimal arrangement of the new scouts into tw patrols By using anonymous data from scouts and parents, a brute force strategy was employed to find an optimal arrangement of scouts that avoids severe conflict while maximizing pre-existing preferences.
 \end{abstract}
@@
 C Physical Review C 65.3 (2002)}.
-Additionally, this model could be applied to other situations in which arranging a number of individuals into groups where interpersonal dynamics are necessary to consider - such as teams within a company or military troops.
+Additionally, this model can be applied to elementary school students to promote supportive and enthusiastic groups for better learning, ensuring cooperation and harmony in military troops for mission success and survival (\textit{Personal Interviews}), and other situations where interpersonal dynamics are necessary.
+
 The 13 different recruited scouts are Brandon, Cameron, Christian, Colby, Daniel, Darwin, Evan, Jake, Jordan, Nathan, Patrick, Timmy and Tommy.
@@
 Ultimately, our goal is to find groups for the scouts that will allow them to enjoy their time scouting to the fullest.
-Modeling interpersonal relations by likes and dislikes is, of couse, a simplification.
+Modeling interpersonal relations by likes and dislikes is, of course, a simplification.
 But the difficulty in modeling the intricacies of human relationships demands it.
@@
 Another simplification is to assume that the quality of a group can be determined by the quality of relationships between unique pairs either inside or outside the group - in other words, the whole is equal to the sum of the parts.
-For example, consider the case where scouts A, B and C individually like each other when they are alone with one of their group, but don't like being in a group together.
-This type of epiphenomenon won't be considered.
+For example, consider the case where scouts A, B and C individually like each other when they are alone with one of their group, but do not like being in a group together.
+This type of epiphenomenon will not be considered.
 
@@
 The total number of partitions given these constrants is $\binom{13}{6} = \binom{13}{7} = 1716$.
-This number is reasonable enough that we may calculate the total loss of every valid grouping and select the partition with the lowest or highest score depending on the objective functionn.
+This number is reasonable enough that we may calculate the total loss of every valid grouping and select the partition with the lowest or highest score depending on the objective function.
 
@@
 
-Of the objective functions we chose to use, it seems like MinCut and HybridCut are the best performing.
+Of the objective functions we chose to use, it seems as though MinCut and HybridCut are the best performing.
 FriendCut performs almost as well as HybridCut, except for one worse in the EnemyCut objective.
@@
 	\item As mentioned in the previous point, Tommy dislikes Brandon.
-Trading Tommy in $P1$ for Jordan in $P2$ seems promising, But we have overlooked the fact that Colby and Jordan are our NA pair, and cannot be in the same patrol under any circumstances.
+Trading Tommy in $P1$ for Jordan in $P2$ seems promising, but we have overlooked the fact that Colby and Jordan are our NA pair, and cannot be in the same patrol under any circumstances.
 Trading Colby for Jordan is another option, but Colby will be missed by Darwin, Evan, Cameron and Tommy (In other words, the whole of P1 minus Brandon).
@@
 / (10!)^{15}$ possible partitions, a number proportional to $10^{164}$ - meaning we could instead use that computational time to enumerate the number of atoms in the universe\ldots a tredecillion $(10^{78})$ times.
-Knowing this, perhaps it won't surprise the reader that this particular graph partitioning problem in general graphs is NP-complete - though there do exist approximation algorithms\footnote{B.
+Knowing this, perhaps it will not surprise the reader that this particular graph partitioning problem in general graphs is NP-complete - though there do exist approximation algorithms\footnote{B.
 W.
@@
 When the group is small enough - as it is here - globally optimal solutions may be arrived at by checking all feasible solutions.
-But the amount of computation time needed to arrive at an answer grows exponentially with the number of individuals and groups.
+However, the amount of computation time needed to arrive at an answer grows exponentially with the number of individuals and groups.
 For example, finding an optimal partition of 13 scouts into \textit{three} groups would take over 50 times as long as finding the optimal partition of 13 scouts into two groups (90,090 feasible combinations to check versus 1,716 in the two patrol case).
@@
 
-\bibitem {B} Desquelles, P. "Calculation of the Number of Partitions with Constraints on the Fragment Size." Phys. Rev. C Physical Review C 65.3 (2002)
+\bibitem {B} Desquelles, P. ``Calculation of the Number of Partitions with Constraints on the Fragment Size." Phys. Rev. C Physical Review C 65.3 (2002)
 
 \bibitem {C} Hartigan, J. A., and M. A. Wong. "Algorithm AS 136: A K-Means Clustering Algorithm." Applied Statistics 28.1 (1979): 100.
+
+\bibitem {D} Mendoza, Veronica, ``Mabel Paine Elementary School classroom helper." \textit{Personal Communication}.
+
+\bibitem {E} Lin, Kevin. ``Air Force ROTC Cadet in the 910th Cadet Wing." \textit{Personal Communication}.
 

### 2016-06-08 draft1.tex
Margin change and more edits

@@
 \usepackage{array}
+
+\setlength{\textwidth}{\paperwidth}
+\addtolength{\textwidth}{-2in}
+\calclayout
 
@@
 For 13 scouts, a grouping into one patrol is too large, while three patrols are too small, and need not be considered.
-But in general it is possible to calculate all possible partitions with size constraints on the partitions for an arbitrary number of scouts\footnote{Desquelles, P.
-"Calculation of the Number of Partitions with Constraints on the Fragment Size." Phys.
-Rev.
-C Physical Review C 65.3 (2002)}.
-Additionally, this model can be applied to elementary school students to promote supportive and enthusiastic groups for better learning, ensuring cooperation and harmony in military troops for mission success and survival (\textit{Personal Interviews}), and other situations where interpersonal dynamics are necessary.
+But in general it is possible to calculate all possible partitions with size constraints on the partitions for an arbitrary number of scouts\cite{B}.
+
+Furthermore, this model can be applied to other situations where interpersonal dynamics is important.
+For example, elementary school students usually are arranged into groups during class activities.
+The kids that formed supportive groups tend to have more enthusiastic students, enhancing their enjoyment and learning during the lesson \cite{D}.
+Another example is military personnel that conduct missions in troops.
+The harmony and cooperation within a troop is necessary to form trust for the success of their mission and, at times, survival of the entire group \cite{E}.
 
@@
 
-The Stirling number of the second kind provides the total possible ways to arrange any number of distinguishable objects to indistinguishable groups as long as any troop size is permitted (except the empty troop).
-For the data in this project, the Stirling number of the second kind for 13 scouts into 2 troops is 4095\footnote{See Appendix B}.
+The Stirling number of the second kind provides the total possible ways to arrange any number of distinguishable objects to indistinguishable groups as long as any troop size is permitted (except the empty troop) \cite{F}.
+For the data in this project, the Stirling number of the second kind for 13 scouts into 2 troops is 4095 (see \textit{Appendix B}).
 This number provides an upper bound on the possible arrangements for the specific problem tackled in this project.
@@
 To model the relationships between scouts, we use a mathematical graph.
-We let the scouts be vertices in our graph.
+A graph is a convenient way to illustrate direct relationships between pairs of objects; for example, if we have a collection of cities in a state, a city can be related to another if there is a direct path between them.
+A city can be considered a vertex, and two cities can have a edge between them if there is a direct path connecting them.
+Edges can have weights attached to them such as the distance between the two cities.
+The edges can even be directed in which the distance from City A to City B might be different with the distance from City B to City A.
+Graphs can also be complete if each vertex in a graph is connected to every other vertex.
+Graphs are not limited to just cities on a map, but rather a collection of objects with any notion of relationship can be represented as a graph.
+
+For our problem, we let the scouts be vertices in our graph.
 A directed edge between Scout A and Scout B has a weight proportional to how much Scout A likes Scout B.
-We would then like to find a partition of the vertices in the graph that either minimizes or maximizes an objective function given constraints on the sizes of subsets generated by our partition.
+A partition of a graph is a grouping of vertices, thus a partition of scouts is considered a troop.
+Whenever our graph is arranged into two partitions, we then can assess this arrangement based on a scoring criterion (or an objective function).
+We would like to find a partition of the vertices in the graph that either minimizes or maximizes an objective function given constraints on the sizes of subsets generated by our partition.
 This results in a complete, directed graph on 13 vertices having edge weights 0, 1, or -1, corresponding to indifference, approval, or disapproval, respectively.
@@
 
-Let's take a more critical look at the optimal partition arrived at via MinCut:
+Let us take a more critical look at the optimal partition arrived at via MinCut:
 
@@
 / (10!)^{15}$ possible partitions, a number proportional to $10^{164}$ - meaning we could instead use that computational time to enumerate the number of atoms in the universe\ldots a tredecillion $(10^{78})$ times.
-Knowing this, perhaps it will not surprise the reader that this particular graph partitioning problem in general graphs is NP-complete - though there do exist approximation algorithms\footnote{B.
-W.
-Kernighan, S.
-Lin, \textit{An Efficient Heuristic Procedure for Partitioning Graphs}.
-Bell System Technical Journal.
-\textbf{49} (1970), 291--307.
-doi: 10.1002/j.1538-7305.1970.tb01770.x}.
+Knowing this, perhaps it will not surprise the reader that this particular graph partitioning problem in general graphs is NP-complete - though there do exist approximation algorithms \cite{A}.
 As scout enrollment increases, it becomes necessary to consider scenarios where we have both a large number of scouts in need of assignment and multiple patrols to choose from.
@@
 
+\bibitem {F} Roberts, F., Tesman, B. ``Applied Combinatorics Second Edition." \textit{CRC Press}. 2009. Pp 54-55.
+
 \end{thebibliography}

### 2016-06-08 draft1.tex

@@
 To model the relationships between scouts, we use a mathematical graph.
-A graph is a convenient way to illustrate direct relationships between pairs of objects; for example, if we have a collection of cities in a state, a city can be related to another if there is a direct path between them.
-A city can be considered a vertex, and two cities can have a edge between them if there is a direct path connecting them.
-Edges can have weights attached to them such as the distance between the two cities.
-The edges can even be directed in which the distance from City A to City B might be different with the distance from City B to City A.
+A graph is a convenient way to illustrate direct relationships between pairs of objects; for example, if we have a collection of cities in a state, a city can be related to another if there is a road between them.
+A city can be considered a vertex, and two cities can have a edge between them if there is a road connecting them.
+Edges can have weights attached to them such as the length of the road between the two cities.
+The edges can even be directed, meaning that distance from City A to City B might be different with the distance from City B to City A (two one-way roads may have different lengths).
 Graphs can also be complete if each vertex in a graph is connected to every other vertex.
@@
 \section*{Verification Statement}
-(Once we receive our verification statement, we will place it here.)
-
+Scoutmaster Gene Bruce addressed our professor, Sara Billey, with the following comments: ``I'd like to thank you and your students for developing a process to divide a group of new scouts into patrols based upon solicited personal preferences.
+While a relatively small problem, it can take a lot of effort to solve by hand.
+The students grasped the subtleties of the conflicting objectives.
+I appreciated their approach of using multiple objective functions.
+This provided a family of solutions that gave insights into the problem.
+I appreciate this much more than a process that produces the ``one best answer".
+It's much easier for me to add my ``gut instinct" with a handful of alternatives rather than the original 1716 possible combinations.
+I look forward to setting up the process on my home computer and leveraging it during the next recruiting season."
 \section*{Appendix A}

### 2016-06-08 draft1.tex
Added Appendix C on counting

--- 2016-06-08 draft1.tex
@@
 This results in a complete, directed graph on 13 vertices having edge weights 0, 1, or -1, corresponding to indifference, approval, or disapproval, respectively.
+
 Scoutmaster Bruce specifically would like two patrols from the 13 scouts - one of size 6, the other of size 7.
-The total number of partitions given these constrants is $\binom{13}{6} = \binom{13}{7} = 1716$.
+Thus, it can be simpler to calculate all possible combinations of the scouts into groups of
+The total number of partitions given these constrants is $\binom{13}{6} = \binom{13}{7} = 1716$ (see \textit{Appendix C}).
 This number is reasonable enough that we may calculate the total loss of every valid grouping and select the partition with the lowest or highest score depending on the objective function.
@@
 
+\section*{Appendix C}
+Scoutmaster Gene Bruce wanted to arrange 13 scouts into two groups where one group will have 6 scouts and the other will have 7 scouts.
+We can find the number of combinations possible of the 13 scouts into a troop size of 6 or 7 using the binomial coefficient.
+
+Let us first tackle a simpler problem to understand how to find the number of combinations.
+Suppose we have 4 people: A, B, C and D that we want to form groups of size 3 where the order of people does not matter.
+Then here are all the possible groups:
+\begin{itemize}
+	\item A,B,C	
+	\item A,B,D
+	\item A,C,D	
+	\item B,C,D
+\end{itemize}
+Thus, there are 4 different combinations of grouping 4 people in a group of size 3.
+It would be impractical to list out all the possible combinations for larger number of people, thus we use the binomial coefficient formula \cite{F}:
+\begin{gather*}
+	\binom{n}{r} = \frac{n!}{r!(n-r)!}
+\end{gather*}
+where $n$ is the number of people and $r$ is the size of the group. For this small problem:
+\begin{gather*}
+	\binom{4}{3} = \frac{4!}{3!(4-3)!} = \frac{4\cdot3!}{3!1!} = 4
+\end{gather*} 
+Thus, we can see we arrive with the same result.
+
+Now to address the question of arranging 13 scouts into a troop of size 6:
+\begin{gather*}
+	\binom{13}{6} = \frac{13!}{6!(13-6)!} = 13!/(6!7!) = 1716
+\end{gather*}
+The other troop would have a size of 7:
+\begin{gather*}
+	\binom{13}{7} = \frac{13!}{7!(13-7)!} = 13!/(7!6!) = 1716
+\end{gather*}
+Therefore the number of combinations possible for arranging 13 scouts into 6 or 7 troop size is 1716.
 \bibliographystyle{amsplain}

### 2016-06-08 draft1.tex
some edits

--- 2016-06-08 draft1.tex
@@
 To model the relationships between scouts, we use a mathematical graph.
-A graph is a convenient way to illustrate direct relationships between pairs of objects; for example, if we have a collection of cities in a state, a city can be related to another if there is a road between them.
-A city can be considered a vertex, and two cities can have a edge between them if there is a road connecting them.
-Edges can have weights attached to them such as the length of the road between the two cities.
-The edges can even be directed, meaning that distance from City A to City B might be different with the distance from City B to City A (two one-way roads may have different lengths).
-Graphs can also be complete if each vertex in a graph is connected to every other vertex.
-Graphs are not limited to just cities on a map, but rather a collection of objects with any notion of relationship can be represented as a graph.
+A graph is a convenient way to illustrate direct relationships between pairs of objects; and a good example that explains how they work is to consider roads between cities.
+On a map, we consider two cities to be ``connected" if they share a road between them; by this, we mean that the existence of a road defines the property of two cities being connected.
+Graphs abstract this notion; instead of considering connections between a map of cities, we consider connections between groups of any kind of object.
+In graph theoretic terminology, we call these objects vertices, and the entity that connects two vertices is called an edge.
+We could redefine our map of cities in this terminology: consider cities as being vertices in a graph where the criterion for having an edge between two vertices is when there is a road built between the cities.
+Another common optional property of graphs is where we ascribe numbers to the edges, corresponding to particular ``weights" that they have.
+In our city graph example, we could consider each edge to have a weight of how long the road is.
+Edges can even be directed, meaning that a particular relationship between vertices is only valid from one vertex to another, and not the other way around.
+For example, there could be a one-way road from city A to city B, and no road from city B to city A.
+In our situation, we will also consider graphs that are complete, where completeness means that each vertex is connected to every other vertex.
 
-For our problem, we let the scouts be vertices in our graph.
+For our problem, we can define a graph where scouts are vertices.
 A directed edge between Scout A and Scout B has a weight proportional to how much Scout A likes Scout B.
-A partition of a graph is a grouping of vertices, thus a partition of scouts is considered a troop.
-Whenever our graph is arranged into two partitions, we then can assess this arrangement based on a scoring criterion (or an objective function).
-We would like to find a partition of the vertices in the graph that either minimizes or maximizes an objective function given constraints on the sizes of subsets generated by our partition.
-This results in a complete, directed graph on 13 vertices having edge weights 0, 1, or -1, corresponding to indifference, approval, or disapproval, respectively.
-Scoutmaster Bruce specifically would like two patrols from the 13 scouts - one of size 6, the other of size 7.
-The total number of partitions given these constrants is $\binom{13}{6} = \binom{13}{7} = 1716$.
-This number is reasonable enough that we may calculate the total loss of every valid grouping and select the partition with the lowest or highest score depending on the objective function.
+We give the directed edge between scout A and scout B a weighting of 1 if A likes B, 0 if A is indifferent to B, and a -1 if A disapproves of B.
+This results in a complete, directed graph of 13 vertices, since there are 13 scouts.
+Note that these edges are directed; we want to consider cases where the relationship between two scouts is not symmetric, whereas a weighted undirected graph would not be able to represent this asymmetry.
+Let us now define a $k$-partition of a graph as a grouping of vertices where each vertex is assigned into one of $k$ groups where $k$ is a natural number; thus, a $k$-partition of the vertices in the scout graph would represent splitting up the scouts into $k$ different patrols.
+Scoutmaster Bruce specifically would like two patrols from the 13 scouts - one of size 6, the other of size 7, which means that we need to find a 2-partition of the graph where 6 vertices are in one grouping and 7 are in the other.
+The total number of partitions given these constraints is $\binom{13}{6} = \binom{13}{7} = 1716$.
+Given these 1716 partitions, we want to select one of them that is ``best" for the scouts in question.
+In mathematics, picking the ``best" option out of a list of options corresponds to first defining a function that ``scores" how good an item is (or bad an item is) and then picking the item with the highest score (or the least score, if our scoring function defines how bad an item is).
+This process of selecting the item from a list that has the most extreme score is called optimization.
+From now on, we will call these scoring functions objective functions.
+Our process for optimizing a partition of the scouts will become computing an objective function for each possible partition and then selecting the partition that gives us the most extreme score.
+This ``brute force" approach to computing the scores isn't always appropriate for optimizing an objective function over a set of values, as the size of the set of values is too large for a computer to iterate over in a timely fashion.
+Often, a more sophisticated technique will be required to iterate over a manageable subset of the set of possible values.
+However, 1716 is within the range of sizes of sets to iterate over where we could apply a brute force technique to solve for the most optimal partition.
 
-We measure the goodness (or score) of a partition in six different ways, corresponding to six different objective functions, which we dub MinCut, FriendCut, EnemyCut, AwkwardCut, and HybridCut.
+We provide six different objective functions that we deemed reasonable for measuring how well patrol partitions would work.
+We dub each of these functions MinCut, FriendCut, EnemyCut, AwkwardCut (+1), AwkwardCut (+2), and HybridCut.
 MinCut is designed to minimize the cut edge weights between partitions (see \textit{Solution of Mathematical Problem}).

### 2016-06-08 main.py

@@ -1,5 +1,3 @@
-# Julio's here!!!
-
 from __future__ import print_function
 import pandas as pd
 import numpy as np

### 2016-06-08 draft1.tex

@@
 \usepackage{array}
-
 \setlength{\textwidth}{\paperwidth}
@@
 
+\begin{lstlisting}
+from __future__ import print_function
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+import itertools
+
+data = pd.read_csv("data.csv", header=None, dtype='float')
+data = data.fillna(float('-inf'))
+names = pd.read_csv("names.csv", header=None)
+
+'''
+Traditional Min Graph Cut
+
+Change in loss function for scouts in *different* patrols
+
+mutual friends: +2
+one like, one indifferent: +1
+one like, one dislike: 0
+one dislike, one indifferent: -1 
+mutual enmity: -2
+'''
+def min_cut(pair, in_same_patrol):
+	x, y = pair
+	if not in_same_patrol:
+		return max(-2, data[x][y] + data[y][x])
+	else:
+		return 0
+
+'''
+Friendly Cut: maximize friendship within patrol
+
+Change in loss function for scouts in *same* patrol:
+
+mutual friends: -2
+one like, one indifferent: -1
+one like, one dislike: 0
+one dislike, one indifferent: 0 
+mutual enmity: 0
+'''
+def friendly_cut(pair, in_same_patrol):
+	x, y = pair
+	if in_same_patrol:
+		return -1 * max(0, data[x][y] + data[y][x])
+	else:
+		return 0
+
+'''
+Enemy Cut: minimize enemies within patrol
+
+Change in loss function for scouts in *same* patrol:
+
+mutual friends:0 
+one like, one indifferent: 0
+one like, one dislike: 0
+one dislike, one indifferent: +1 
+mutual enmity: +2 
+'''
+def enemy_cut(pair, in_same_patrol):
+	x, y = pair
+	if in_same_patrol:
+		return max(0, -(data[x][y] + data[y][x]))
+	else:
+		return 0
+
+'''
+Awkward Cut: same as Min Cut but with additional penalty
+for one like, one dislike in same patrol
+
+Change in loss function:
+
+mutual friends: +2
+one like, one indifferent: +1
+one like, one dislike (same patrol): +1
+one like, one dislike (different patrol): 0
+one dislike, one indifferent: -1 
+mutual enmity: -2
+'''
+def awkward_cut(pair, in_same_patrol):
+	x, y = pair
+	if in_same_patrol and data[x][y] == -data[y][x] and abs(data[x][y]) > 0:
+		return 1 
+	elif not in_same_patrol:
+		return max(-2, data[x][y] + data[y][x])
+	else:
+		return 0
+
+def hybrid_cut(pair, in_same_patrol):
+	return 0.5 * min_cut(pair, in_same_patrol) \
+		+ 0.5 * friendly_cut(pair, in_same_patrol)
+
+# this function assumes that the partitions are disjoint, no
+# check for this has been implemented
+def calculate_loss(partition, objective):
+	def compare_pair(pair): # returns true if pair is in the same partition,
+		ans = False         # false otherwise
+		for part in partition:
+			ans = ans or (pair[0] in part and pair[1] in part)
+		return ans
+	total = ()			# total will include every element in the
+	loss = 0			# partitions
+	for part in partition:
+		total += tuple(part)
+	every_pair = itertools.combinations(total, 2)
+	for pair in every_pair:
+		loss += objective(pair, compare_pair(pair))
+	return loss
+
+def brute_force_solver(d=data, num_patrols=2, objective=min_cut, min_patrol_size=6, max_patrol_size=7):
+	partitions = generate_partitions([], num_patrols, len(d))
+	best_partition_so_far = None
+	best_loss = float('inf')
+	for part in partitions:
+		if (sum(part) >= min_patrol_size and sum(part) <= max_patrol_size \
+			and not (part[3] and part[8]) and not(not(part[3] or part[8]))):
+			if any([3 in p and 8 in p for p in partitions]): print('hello')
+			patrols = [[] for i in range(num_patrols)]
+			for i in range(len(part)):
+				patrols[part[i]] += [i]
+			loss = calculate_loss(patrols, objective)	
+			if loss < best_loss:
+				best_partition_so_far = patrols
+				best_loss = loss
+	return best_partition_so_far, best_loss
+
+def generate_partitions(arr, patrols, num_scouts):
+	result = []
+	if num_scouts == 0:
+		return [arr]
+	else:
+		for i in range(patrols):
+			for c in generate_partitions(arr + [i], patrols, num_scouts - 1):
+				result += [c]
+	return result
+'''
+Example use
+
+best_patrol, best_loss = brute_force_solver()
+print "Best patrol:", best_patrol
+print "Loss:", best_loss
+'''
+\end{lstlisting}
 \end{thebibliography}

### 2016-06-08 draft1.tex
merge
formatting
footnote
Add appendix C in text

@@
 \usepackage{array}
+\usepackage{listings}
 \setlength{\textwidth}{\paperwidth}
-\addtolength{\textwidth}{-2in}
+\addtolength{\textwidth}{-3in}
 \calclayout
@@
 Scoutmaster Bruce specifically would like two patrols from the 13 scouts - one of size 6, the other of size 7, which means that we need to find a 2-partition of the graph where 6 vertices are in one grouping and 7 are in the other.
-The total number of partitions given these constraints is $\binom{13}{6} = \binom{13}{7} = 1716$.
+The total number of partitions given these constraints is $\binom{13}{6} = \binom{13}{7} = 1716$ (see \textit{Appendix C}).
 Given these 1716 partitions, we want to select one of them that is ``best" for the scouts in question.
@@
 
-\clearpage
-
 For each method we define three scoring metrics:
@@
 \end{figure}
-
-\clearpage
 
@@
 Nevertheless, our chosen objective functions seem to arrive at solutions that are able to hold up to human scrutiny, and perform well enough when total enumeration is possible.
-Larger problems may require approximating or clustering methods\footnote{Hartigan, J.
-A., and M.
-A.
-Wong.
-"Algorithm AS 136: A K-Means Clustering Algorithm." Applied Statistics 28.1 (1979): 100.}.
+Larger problems may require approximating or clustering methods[3].
 
@@
 
+\clearpage
+
 \section*{Appendix C}
@@
 Suppose we have 4 people: A, B, C and D that we want to form groups of size 3 where the order of people does not matter.
-Then here are all the possible groups:
+The possible combinations are:
+
 \begin{itemize}
@@
 \bibitem {F} Roberts, F., Tesman, B. ``Applied Combinatorics Second Edition." \textit{CRC Press}. 2009. Pp 54-55.
+
+\clearpage
+
+\section*{Code}
 

### 2016-06-08 draft1.tex
final amendments

@@
 For 13 scouts, a grouping into one patrol is too large, while three patrols are too small, and need not be considered.
-But in general it is possible to calculate all possible partitions with size constraints on the partitions for an arbitrary number of scouts\cite{B}.
+But in general it is possible to calculate all possible partitions with size constraints on the partitions for an arbitrary number of scouts\cite{A}.
 
@@
 For example, elementary school students usually are arranged into groups during class activities.
-The kids that formed supportive groups tend to have more enthusiastic students, enhancing their enjoyment and learning during the lesson \cite{D}.
+The kids that formed supportive groups tend to have more enthusiastic students, enhancing their enjoyment and learning during the lesson \cite{B}.
 Another example is military personnel that conduct missions in troops.
-The harmony and cooperation within a troop is necessary to form trust for the success of their mission and, at times, survival of the entire group \cite{E}.
+The harmony and cooperation within a troop is necessary to form trust for the success of their mission and, at times, survival of the entire group \cite{C}.
 
@@
 
-The Stirling number of the second kind provides the total possible ways to arrange any number of distinguishable objects to indistinguishable groups as long as any troop size is permitted (except the empty troop) \cite{F}.
+The Stirling number of the second kind provides the total possible ways to arrange any number of distinguishable objects to indistinguishable groups as long as any troop size is permitted (except the empty troop) \cite{D}.
 For the data in this project, the Stirling number of the second kind for 13 scouts into 2 troops is 4095 (see \textit{Appendix B}).
@@
 Scoutmaster Bruce specifically would like two patrols from the 13 scouts - one of size 6, the other of size 7, which means that we need to find a 2-partition of the graph where 6 vertices are in one grouping and 7 are in the other.
-The total number of partitions given these constraints is $\binom{13}{6} = \binom{13}{7} = 1716$ (see \textit{Appendix C}).
+The total number of partitions given these constraints is $\binom{13}{6} = \binom{13}{7} = 1716$.
 Given these 1716 partitions, we want to select one of them that is ``best" for the scouts in question.
@@
 Our process for optimizing a partition of the scouts will become computing an objective function for each possible partition and then selecting the partition that gives us the most extreme score.
-This ``brute force" approach to computing the scores isn't always appropriate for optimizing an objective function over a set of values, as the size of the set of values is too large for a computer to iterate over in a timely fashion.
+This ``brute force" approach to computing the scores isn't always appropriate for optimizing an objective function over a set of values, as the size of the set of values is often too large for a computer to iterate over in a timely fashion.
 Often, a more sophisticated technique will be required to iterate over a manageable subset of the set of possible values.
@@
 / (10!)^{15}$ possible partitions, a number proportional to $10^{164}$ - meaning we could instead use that computational time to enumerate the number of atoms in the universe\ldots a tredecillion $(10^{78})$ times.
-Knowing this, perhaps it will not surprise the reader that this particular graph partitioning problem in general graphs is NP-complete - though there do exist approximation algorithms \cite{A}.
+Knowing this, perhaps it will not surprise the reader that this particular graph partitioning problem in general graphs is NP-complete - though there do exist approximation algorithms \cite{E}.
 As scout enrollment increases, it becomes necessary to consider scenarios where we have both a large number of scouts in need of assignment and multiple patrols to choose from.
@@
 Nevertheless, our chosen objective functions seem to arrive at solutions that are able to hold up to human scrutiny, and perform well enough when total enumeration is possible.
-Larger problems may require approximating or clustering methods[3].
+Larger problems may require approximating or clustering methods\cite{F}.
 
@@
 
-\bibitem {A} B.
+\bibitem {A} Desquelles, P. ``Calculation of the Number of Partitions with Constraints on the Fragment Size." Phys. Rev. C Physical Review C 65.3 (2002)
+
+\bibitem {B} Mendoza, Veronica, ``Mabel Paine Elementary School classroom helper." \textit{Personal Communication}.
+
+\bibitem {C} Lin, Kevin. ``Air Force ROTC Cadet in the 910th Cadet Wing." \textit{Personal Communication}.
+
+\bibitem {D} Roberts, F., Tesman, B. ``Applied Combinatorics Second Edition." \textit{CRC Press}. 2009. Pp 54-55.
+
+\bibitem {E} B.
 W.
@@
 
-\bibitem {B} Desquelles, P. ``Calculation of the Number of Partitions with Constraints on the Fragment Size." Phys. Rev. C Physical Review C 65.3 (2002)
-
-\bibitem {C} Hartigan, J. A., and M. A. Wong. "Algorithm AS 136: A K-Means Clustering Algorithm." Applied Statistics 28.1 (1979): 100.
-
-\bibitem {D} Mendoza, Veronica, ``Mabel Paine Elementary School classroom helper." \textit{Personal Communication}.
-
-\bibitem {E} Lin, Kevin. ``Air Force ROTC Cadet in the 910th Cadet Wing." \textit{Personal Communication}.
-
-\bibitem {F} Roberts, F., Tesman, B. ``Applied Combinatorics Second Edition." \textit{CRC Press}. 2009. Pp 54-55.
+\bibitem {F} Hartigan, J. A., and M. A. Wong. "Algorithm AS 136: A K-Means Clustering Algorithm." Applied Statistics 28.1 (1979): 100.
 

### 2016-06-08 draft1.tex

@@
 Our process for optimizing a partition of the scouts will become computing an objective function for each possible partition and then selecting the partition that gives us the most extreme score.
-This ``brute force" approach to computing the scores isn't always appropriate for optimizing an objective function over a set of values, as the size of the set of values is often too large for a computer to iterate over in a timely fashion.
+This ``brute force" approach to computing the scores isn't always appropriate for optimizing an objective function over a set of values, as the size of the set of values is too large for a computer to iterate over in a timely fashion.
 Often, a more sophisticated technique will be required to iterate over a manageable subset of the set of possible values.
